---
title: "Project Study"
author: "Helena Zappe & Marei Göbelbecker"
date: "2025-06-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Download Packages

```{r install-packages, eval=FALSE, include=FALSE}
install.packages(c(
  "tidyverse",  # Datenaufbereitung & Visualisierung
  "skimr",      # Übersichtliche Daten-Zusammenfassung
  "psych",      # Deskriptive Statistik
  "rmarkdown",  # Für R Markdown selbst
  "knitr",      # Code-Ausgabe in R Markdown
  "tinytex",    # Nur für PDF-Export nötig
  "labelled"
  ))
tinytex::install_tinytex()  # Nur beim ersten Mal notwendig
```

# Load Packages

```{r load-packages, include=FALSE}
library(tidyverse)   # Bündelt ggplot2, dplyr, readr, etc. – für Datenimport, -aufbereitung, Visualisierung
library(skimr)       # Für übersichtliche Zusammenfassungen von Datensätzen
library(psych)       # Für deskriptive Statistiken (z.B. Mittelwert, SD, Median)
library(rmarkdown)   # Für das Erstellen von HTML/PDF-Dokumenten
library(knitr)       # Macht Code-Ausgabe in R Markdown hübsch und steuerbar
library(tinytex)     # Wird nur gebraucht, wenn du PDF-Dateien erstellst
library(labelled)
library(dplyr)
library(ggplot2)
library(tidyr)
library(forcats)
```

# Loading data

```{r}

setwd("/Users/test/Documents/Studium/Semester 6/Project Study")

dt_raw <- haven::read_sav(file.choose()) #ESS_all.sav

```

*`"What factors shapes trust in the European Union in Poland, Germany, and Slovenia? A comparative analysis of political attitudes, ideology, and socio-demographics."`*

# Data Wrangling

-   filtering to our countries
-   recoding age group
-   reversing the scale for polintr (higher number = more interested)
-   adjusting labels for ESS rounds
-   winsorizing education years to get rid of outliers

```{r data_preparation}

dt_all <- dt_raw %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%  
  mutate(
    age_group = case_when(
      agea < 18 ~ "<18",                      
      agea >=18 & agea <= 24 ~ "18-24",
      agea > 24 & agea <= 34 ~ "25-34",
      agea > 34 & agea <= 44 ~ "35-44",
      agea > 44 & agea <= 54 ~ "45–54",
      agea > 54 & agea <= 64 ~ "55-64",
      agea > 64 ~ "65+"
    ))  %>%                                   
  mutate(
    age_group = factor(age_group, levels = c("<18", "18-24", "25-34", "35-44", "45–54", "55-64", "65+")),
    polintr_rev = 5 - polintr  # higher = more interested               
    ) %>%
  mutate(year = case_when(    # Mapping ESS round to the year they were conducted in in "year"
    essround == 1 ~ 2002,
    essround == 2 ~ 2004,
    essround == 3 ~ 2006,
    essround == 4 ~ 2008,
    essround == 5 ~ 2010,
    essround == 6 ~ 2012,
    essround == 7 ~ 2014,
    essround == 8 ~ 2016,
    essround == 9 ~ 2018,
    essround == 10 ~ 2020,
    essround == 11 ~ 2022,
    TRUE ~ NA_real_
  )) %>%
    select(-c(agea, polintr, essround, trstun, trstplc, trstplt, trstprl, trstprt, name, pplfair, pplhlp, stfdem, imueclt, imdfetn, imbgeco, impcntr, badge))

```

```{r}
# winsorize education years to get rid of outliers
# Calculate thresholds
q_lower <- quantile(dt_all$eduyrs, 0.01, na.rm = TRUE)
q_upper <- quantile(dt_all$eduyrs, 0.99, na.rm = TRUE)

# Winsorize: replace values outside bounds
dt_all <- dt_all %>%
  mutate(eduyrs_winsor = case_when(
    eduyrs < q_lower ~ q_lower,
    eduyrs > q_upper ~ q_upper,
    TRUE ~ eduyrs
  )) %>%
select(-eduyrs)

```

```{r}
# Saving Data as new Dataset for Lasso

write_sav(dt_all, "ESS_prepared.sav") 

```


```{r raw_data, include=FALSE}

#Lade Daten und entferne nicht benötigte Variablen
dt_all <- haven::read_sav(file.choose()) 

#ESS_prepared.sav 

```

# LASSO to find best variables

```{r}

# Step 1: Preparing Data for Lasso
# using the gilmnet package
  
# filter variables with too many NAs:
dt_filtered <- dt_all %>%
  drop_na(trstep) %>% # keep only valid observations in y
  select(where(~ !all(is.na(.)))) %>% # 100% NAs
  select(where(~ mean(is.na(.)) <= 0.5)) %>% # keep variables with less than 80% NAs
 # select(-c(feethngr, loylead, lrnobed, donprty, volunfp, pbldmna, implvdm)) %>% # these variables are not relevant enough for imputation. Also households net income has 79% NAs and is therefore not meaningful. 
  select(-c(idno, proddate, bctprd, wrkorg, edition)) # Other variables that are not relevant for interpretation, stratum

# checking for missing variables in percent
dt_filtered %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_pct") %>%
  filter(missing_pct > 20) %>%
  arrange(desc(missing_pct))

# accounting for differences in country size and selection
dt_filtered <- dt_filtered %>% drop_na(anweight) # dropping all obs where anweight is NA
weights <- dt_filtered$anweight # establishing the weighting

# defining the response Variable
y <- as.numeric(as_factor(dt_filtered$trstep))
table(y, useNA = "always")

# define the predictor variable matrix
x_data <- dt_filtered %>%
  select(-c(trstep, anweight, pspwght, dweight, pweight)) %>% #prob
  zap_labels() %>%  # delete all labels
  mutate(across(where(is.factor), ~ as.numeric(as.character(.)))) # define all factor variables also as numeric

x_data[is.na(x_data)] <- 0  # replaces all NA in x_data

# define the matrix of response variables
x <- data.matrix(x_data)

# As long as you never change anything in dt_all (and you don't re‐run sample() anywhere else),
# calling cv.glmnet() right after this will always split into the same 10 folds.
set.seed(123) 

# LASSO with Crossvalidation and weighting

lasso_cv <- cv.glmnet(x, y, alpha = 1, weights = weights)

```

## Fitting the LASSO Model

The bigger lambda, the more coefficients will be set to 0 for variable selection.

**best_lambda** (lambda.min) gives the most precise model:

```{r}
# Find best Lambda
best_lambda <- lasso_cv$lambda.min

# Final LASSO Model
lasso_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)

# Lambda Precise: 0.001855

# Extracting the Coefficients
coef_lasso <- coef(lasso_model)

selected_vars <- rownames(coef_lasso)[which(coef_lasso[, 1] != 0)]
selected_vars <- selected_vars[selected_vars != "(Intercept)"]

print(selected_vars)

```

**lambda.1se** gives the more robust, more economic (sparsamere) model: (Oft besser generalisierbar)

```{r}
lambda_1se <- lasso_cv$lambda.1se

model_robust <- glmnet(x, y, alpha = 1, lambda = lasso_cv$lambda.1se)

# Lambda Robust: 0.017295

# Extracting the Coefficients
coef_robust <- coef(model_robust)

vars_robust <- rownames(coef_robust)[which(coef_robust[, 1] != 0)]
vars_robust <- vars_robust[vars_robust != "(Intercept)"]

print(vars_robust)
```

Chatty: "In der LASSO-Analyse wurde mit lambda.min = 0.00140 das Modell mit der höchsten Vorhersagegenauigkeit identifiziert. Dieses Modell verwendet relativ viele Prädiktoren.

Alternativ wurde mit lambda.1se = 0.01575 ein robusteres Modell berechnet, das sparsamer ist und potenziell besser generalisiert. Hier bleiben nur die wichtigsten Prädiktoren erhalten.

Je nach Forschungsziel (Präzision vs. Interpretierbarkeit) kann eines der beiden Modelle bevorzugt werden."

## Model Specific Coefficients:

```{r}

# Vergleich
setdiff(selected_vars, vars_robust)  # Variablen nur im präzisen Modell
setdiff(vars_robust, selected_vars)  # Variablen nur im sparsamen Modell

```

```{r}
plot(lasso_cv)
abline(v = log(lasso_cv$lambda.min), col = "blue", lty = 3)
abline(v = log(lasso_cv$lambda.1se), col = "red", lty = 3)
abline(v = log(lambda_max), col = "green", lty = 3)
legend("topright", legend = c("lambda.min", "lambda.1se", "lambda_max"),
       col = c("blue", "red", "green"), lty = 3)
title("Cross-Validation: MSE über Lambda")
```

Chatty: - lambda.min (links) ergibt den niedrigsten mittleren Fehler – aber nutzt mehr Prädiktoren - lambda.1se (weiter rechts) hat leicht höheren Fehler, ist aber viel sparsamer Der Plot zeigt schön, dass der Fehler in einem sehr flachen Bereich minimal ist → daher kannst du das robustere, sparsamere Modell (lambda.1se) gut vertreten.

-\> Wir verwenden also die Variablen vom 2. Modell oder grenzen es weiter manuell ein.

## Neues Datenset lambda_max

```{r}
#dt_lasso_selected <- dt_all %>% select(all_of(vars_robust), trstep)
# insg. 66 variables (zu viele)
```

## Alternative: Interpretierbareres Modell mit max x Variablen

```{r}
lasso_model_max <- glmnet(x, y, alpha = 1, dfmax = 14)

# Extracting the Coefficients
coef_lasso <- coef(lasso_model_max)

# Anzahl ≠ 0-Koeffizienten je Lambda (inkl. Intercept)
nonzero <- lasso_model_max$df
lambda_index <- which(nonzero == 14)[1]  # x Prädiktoren + 1 Intercept

# Wenn gefunden:
if (!is.na(lambda_index)) {
  lambda_max <- lasso_model_max$lambda[lambda_index]
  coef_max <- coef(lasso_model_max, s = lambda_max)
  vars_max <- rownames(coef_max)[which(coef_max[, 1] != 0)]
  vars_max <- vars_max[vars_max != "(Intercept)"]
  print(vars_max)
} else {
  message("Kein Lambda mit genau x Prädiktoren gefunden.")
}
```


### Cutoff ausprobieren:

Chatty empfiehlt 10-30 Variablen

(Je geringer lambda, desto genauer passt das Model.) 

- Lambda 10 vars = not found 
- Lambda 15 vars = 0.1245 
- Lambda 16 vars = 0.1134
- Lambda 17 vars = not found
- Lambda 18 vars = 0.0858
- Lambda 19 vars = not found
- Lambda 20 vars = not found 
- Lambda 25 vars = 0.0591 
- Lambda 30 vars = not found 

Ich glaube 15 oder max. 20 sollten genug sein, sonst ist es zu viel.

### Neues Subset aus lambda_max 

```{r}
dt_lasso_selected <- dt_filtered %>% select(all_of(vars_max), trstep)
# insg. 15 variables

```

### Neues Subset mit originaler Codierung und Labels, aber nur lambda_max variablen

```{r}
#1. Neues Datenset mit Originalstruktur erstellen
  dt_lasso_labeled <- dt_all %>%
    select(all_of(vars_max), trstep)  
  
#2. Nur gültige Fälle von trstep behalten 
  dt_lasso_labeled <- dt_lasso_labeled %>%
    drop_na(trstep)  

#3. Optional: Labels prüfen
  library(labelled)
  look_for(dt_lasso_labeled)
```

### Speichern des Datensets in Dateien

```{r}
write_sav(dt_lasso_labeled, "ESS_lasso.sav")

```

# Descriptive Statistics
## Summary Statistics of Variables by Country


```{r}
summary_table <- dt_lasso_labeled %>%
  group_by(cntry) %>%
  summarise(across(
    c(ppltrst, euftf, freehms, stfedu, stfgov, stfhlth, trstlgl, vote, imwbcnt, health, rlgatnd, gndr, year, trstep),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  )) 

View(as.data.frame(summary_table))

```

-   Mean political interest highest in Germany and lowest in Slovenia
-   Education years highest in Germany and lowest in Slovenia
-   On the left-right scale, Germany is the leftest and Poland the one the furthest right
-   Satisfaction with democracy highest in Germany and lowest in Slovenia
-   Satisfaction with government highest in Germany and lowest in Poland
-   Immigration attitude towards immigrants highest in Poland and lowest in Slovenia
-   Trust in politicians highest in Germany and lowest in Poland, but very low across all countries
-   Trust in EU highest in Poland and lowest in Germany, but only very small differences

## Summary Statistics of Trust in EU by Country

```{r}
dt_lasso_labeled %>%
  group_by(cntry) %>%
  summarise(across(
    c(trstep),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  ))

```

-   Mean trust in EU is rather similar across all three countries

# Visualization


## Trust in EU over Time for all three Countries (Option 1)

```{r}

# Summarize mean trust per country and year

trust_trends <- dt_filtered %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%
  group_by(cntry, year) %>%
  summarise(
    trust_EU = mean(trstep, na.rm = TRUE),
    .groups = "drop"
  ) 

#Reshaping to long format for plotting
trust_long <- trust_trends %>%
  pivot_longer(cols = c(trust_EU),
               names_to = "institution",
               values_to = "trust")

#Plotting the time trend
ggplot(trust_long, aes(x = year, y = trust, color = institution, linetype = institution)) +
  geom_line(size = 1) +
  facet_wrap(~cntry) +
  labs(
    title = "Trust in the EU over Time",
    x = "Year",
    y = "Mean Trust Level",
    color = "Institution",
    linetype = "Institution"
  ) +
  theme_minimal()

```

-   Slovenia shows the highest fluctuations in mean trust in the EU, with a sharp decline between 2006 and 2014

## Trust in EU Time for all three Countries (Option 2)

```{r}

ggplot(trust_trends, aes(x = year, y = trust_EU, color = cntry)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Trust in the EU Over Time",
    x = "Year",
    y = "Mean Trust in EU",
    color = "Country"
  ) +
  theme_minimal()


```

-   strong flucutations over time fro all three countries
-   mean trust in Slovenia peaked in 2006 and was the lowest in 2014
-   mean trust in Poland peaked in 2002 and was the lowest in 2014
-   mean trust in Germany peaked in 2023 and was the lowest in 2010

## Distribution of Age Groups

```{r}

dt_filtered %>%
  count(cntry, age_group) %>%
  group_by(cntry) %>%
  mutate(prop = n / sum(n) * 100) %>%
  ggplot(aes(x = cntry, y = prop, fill = age_group)) +
  geom_col(position = "dodge") +
  labs(
    title = "Relative Distribution of Age Groups by Country",
    x = "Country",
    y = "Proportion of Respondents",
    fill = "Age Group"
  ) +
  theme_minimal()

```

-   Among all countries, the age group "65+" has the highest share
-   Age distribution has an upward trend in all three countries, indicating an aging population

## Average Trust in EU by Age Group and Country

```{r}
dt_filtered %>%
  group_by(cntry, age_group) %>%
  summarise(mean_trustEU = mean(trstep, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_trustEU, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",     # Germany - Blue
      "PL" = "#f6ae2d",     # Poland - Light Orange
      "SI" = "#f26419"      # Slovenia - Dark Orange/Pink
    ),
    labels = c(
      "DE" = "Germany",
      "PL" = "Poland",
      "SI" = "Slovenia"
    )
  ) +
  labs(
    title = "Average Trust in EU by Age Group and Country",
    x = "Age Group",
    y = "Mean Trust in EU",
    fill = "Country"
  ) +
  theme_minimal()

```

-   highest mean trust in EU among people younger than 18 in all three countries, with respondents from Germany having the highest trust
-   general downward trend in meant trust in EU across all countries up to the age group 55-64
-   trust increases again for respondents being 65 or older

## Average Trust in Politicians by Age Group and Country

```{r}
dt_filtered %>%
  group_by(cntry, age_group) %>%
  summarise(mean_trustpol = mean(polintr_rev, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_trustpol, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",     # Germany - Blue
      "PL" = "#f6ae2d",     # Poland - Light Orange
      "SI" = "#f26419"      # Slovenia - Dark Orange/Pink
    ),
    labels = c(
      "DE" = "Germany",
      "PL" = "Poland",
      "SI" = "Slovenia"
    )
  ) +
  labs(
    title = "Average Trust in Politicians by Age Group and Country",
    x = "Age Group",
    y = "Mean Trust in Politicians",
    fill = "Country"
  ) +
  theme_minimal()
```

-   clearly highest trust in politicians in Germany among all age groups
-   Polish respondents seem to have higher trust than Slovenian respondents in all age groups, except for the group 55-64 

## Left-Right Political Placement across Countries

```{r}

# Filtering to valid left-right scale values
lr_data <- dt_filtered %>%
  filter(!is.na(lrscale)) %>%
  group_by(cntry, lrscale) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(lr_data, aes(x = cntry, y = percentage, fill = factor(lrscale, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "Left-Right Scale"
  ) +
  labs(
    title = "Placement on Left-Right Political Scale",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

-   highest share of respondents selected 5 as their political attitude
-   higher share of respondents in Germany and Slovenia are rather left
-   higher share of respondents in Poland are rather right

## Satisfaction with Government by Country

```{r}

ggplot(dt_filtered, aes(x = cntry, y = stfgov, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Satisfaction with Government by Country",
    y = "Satisfaction with Government (Higher = More)",
    x = "Country",
    fill = "Country"
  ) +
  theme_minimal()


```

-   satisfation with democracy highest in Germany and similar in Slovenia and Poland

## Education Years per Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = eduyrs_winsor, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Education Years by Country",
       y = "Years of Education", x = "Country")

```

- we winsorized outliers earlier
- highest years of education in Germany

## Years of Education by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(mean_eduyrs = mean(eduyrs_winsor, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_eduyrs, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Mean Years of Education by Age Group and Country",
    x = "Age Group",
    y = "Mean Years of Education",
    fill = "Country"
  ) +
  theme_minimal()
```

-   graph illustrates differences among age groups across countries
-   Germany has the highest average years of education among all age groups, followed by Poland

## Attitudes towards Immigrants by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(mean_imwbcnt = mean(imwbcnt, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_imwbcnt, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Attitudes towards Immigrants by Age Group and Country",
    x = "Age Group",
    y = "Attitudes towards Immigrants",
    fill = "Country"
  ) +
  theme_minimal()
```

-   Polish respondents seem to be more welcoming of immigrants among all age groups
-   Slovenian respondents seem to be the least welcoming of immigrants among all age groups
-   this seems to be a contradiction to the distribution of the left-right scale, where a higher share of Polish respondents indicated to be more right, and a higher share of Slovenian respondents indicated to be more left
-   a boxplot in the next step could help to check for outliers
-   reasons??

## Immigration Attitude per Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = imwbcnt, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Immigration Attitude by Country",
       y = "Immigration Attitude", x = "Country")
```

-   there do not seem to be high outliers for positive immigration attitude among Polish respondents
-   Why are Polish respondents so welcoming when they are rather right? Would be interesting to see if the respondents who are rather right are also more welcoming, or if the average is high because of the very high immigration attitude among rather left respondents
