---
title: "Project Study"
author: "Marei Göbelbecker & Helena Zappe"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---



# Visualization
## Trust in EU - Dependent Variable
### Trust in EU over Time for all three Countries

```{r}

# Summarize mean trust per country and year
trust_trends <- dt_filtered %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%
  group_by(cntry, year) %>%
  summarise(
    trust_EU = mean(trstep, na.rm = TRUE),
    .groups = "drop"
  ) 

#Reshaping to long format for plotting
trust_long <- trust_trends %>%
  pivot_longer(cols = c(trust_EU),
               names_to = "institution",
               values_to = "trust")

ggplot(trust_trends, aes(x = year, y = trust_EU, color = cntry)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Trust in the EU Over Time",
    x = "Year",
    y = "Mean Trust in EU",
    color = "Country"
  ) +
  theme_minimal()


```

-   strong flucutations over time fro all three countries
-   mean trust in Slovenia peaked in 2006 and was the lowest in 2014
-   mean trust in Poland peaked in 2002 and was the lowest in 2014
-   mean trust in Germany peaked in 2023 and was the lowest in 2010
-   Slovenia shows the highest fluctuations in mean trust in the EU,
    with a sharp decline between 2006 and 2014
    
### Average Trust in EU by Age Group and Country

```{r}
dt_filtered %>%
  group_by(cntry, age_group) %>%
  summarise(mean_trustEU = mean(trstep, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_trustEU, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",     # Germany - Blue
      "PL" = "#f6ae2d",     # Poland - Light Orange
      "SI" = "#f26419"      # Slovenia - Dark Orange/Pink
    ),
    labels = c(
      "DE" = "Germany",
      "PL" = "Poland",
      "SI" = "Slovenia"
    )
  ) +
  labs(
    title = "Average Trust in EU by Age Group and Country",
    x = "Age Group",
    y = "Trust in EU (0-10 Scale)",
    fill = "Country"
  ) +
  theme_minimal()

```

-   highest mean trust in EU among people younger than 18 in all three
    countries, with respondents from Germany having the highest trust
-   general downward trend in meant trust in EU across all countries up
    to the age group 55-64
-   trust increases again for respondents being 65 or older
    
## Socio-Demographics

### Distribution of Gender by Country

```{r}

dt_filtered %>%
  mutate(gndr = as_factor(gndr)) %>%  # convert haven_labelled to factor
  count(cntry, gndr) %>%
  group_by(cntry) %>%
  mutate(prop = n / sum(n) * 100) %>%
  ggplot(aes(x = cntry, y = prop, fill = gndr)) +
  geom_col(position = "dodge") +
    scale_fill_manual(
    values = c(
      "1" = "#33658a",   # male - Blue
      "2" = "#f6ae2d"    # female - Light Orange
    ),
    labels = c("1" = "Male", "2" = "Female")
  ) +
  labs(
    title = "Relative Distribution of Gender by Country",
    x = "Country",
    y = "Proportion of Respondents",
    fill = "Gender"
  ) +
  theme_minimal()

```

-   distribution among all three countries rather similar around 50% for
    each gender
-   Germany has slightly more male respondents
-   Poland and Slovenia have slightly more female respondents
-   gap is the highest in Slovenia


### Distribution of Age Groups

```{r}

dt_filtered %>% 
  count(cntry, age_group) %>% 
  group_by(cntry) %>% 
  mutate(prop = n / sum(n) * 100) %>% 
  ggplot(aes(x = age_group, y = prop, fill = cntry)) +
  geom_col(position = "dodge") +
    scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Age Group Distribution Within Each Country",
    x = "Age Group",
    y = "Proportion of Respondents (%)",
    fill = "Country"
  ) +
  theme_minimal()

```

-   Among all countries, the age group "65+" has the highest share
-   Age distribution has an upward trend in all three countries,
    indicating an aging population

### Years of Education

#### by Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = eduyrs_winsor, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Education Years by Country",
       y = "Years of Education", x = "Country")

```

-   we winsorized outliers earlier
-   highest years of education in Germany

## Political Attitudes
### Distribution of Voting Behavior 

```{r}

# Filtering to valid left-right scale values
vote_data <- dt_filtered %>%
  filter(!is.na(vote)) %>%
  group_by(cntry, vote) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(vote_data, aes(x = cntry, y = percentage, fill = factor(vote, levels = rev(0:3)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "1" = "#002060", "2" = "#4A90E2", "3" = "#DDDDDD"
    ),
    name = "Voted Last National Election"
  ) +
  labs(
    title = "Distribution of Voting Behavior",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

-   highest proportion of voters in Germany
-   Slovenia has the highest share of people that did not vote

### Left-Right Political Placement across Countries

```{r}

# Filtering to valid left-right scale values
lr_data <- dt_filtered %>%
  filter(!is.na(lrscale)) %>%
  group_by(cntry, lrscale) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(lr_data, aes(x = cntry, y = percentage, fill = factor(lrscale, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "Left-Right Scale"
  ) +
  labs(
    title = "Placement on Left-Right Political Scale",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

-   highest share of respondents selected 5 as their political attitude
-   higher share of respondents in Germany and Slovenia are rather left
-   higher share of respondents in Poland are rather right

### Satisfaction with National Government

#### by Country over Time on Average

```{r}
dt_filtered %>%
  filter(!is.na(stfgov), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_stfgov = mean(stfgov, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_stfgov, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
    scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Satisfaction with National Government Over Time",
    x = "Year",
    y = "Satisfaction (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()

```

Interpretation from other graph with distributions:
-   satisfation with national government fluctuated over time among all
    countries
-   Polish respondents were very dissatisfied in 2020
-   German respondents were very dissatisfied in 2002
-   Slovenian respondents were very dissatisfied in 2012

### Satisfaction with State of Education by Country

#### by Country Over Time

```{r}

dt_filtered %>%
  filter(!is.na(stfgov), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_stfedu = mean(stfedu, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_stfedu, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
    scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Satisfaction with State of Education by Country and Year",
    x = "Year",
    y = "Satisfaction (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()

```

-   highest satisfaction with education to be seen in Slovenia, followed
    by Poland and then Germany
    
### Satisfaction with the State of Health within Country

#### by Country over Time on Average

```{r}
dt_filtered %>%
  filter(!is.na(stfhlth), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_stfhlth = mean(stfhlth, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_stfhlth, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
      scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Satisfaction with State of Health by Country over Time",
    x = "Year",
    y = "Satisfaction (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()
```

-   Germany peaked around 2016 just above 6, also the highest value
    among all three countries, but satisfaction declined after that
-   Polish satisfaction crashed to under 3 in 2020
-   values from Polish respondents are the lowest among all three
    countries

### Trust in the Legal System by Country over Time on Average

```{r}

dt_filtered %>%
  filter(!is.na(trstlgl), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_trstlgl = mean(trstlgl, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_trstlgl, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
      scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Trust in the Legal System by Country over Time",
    x = "Year",
    y = "Trust (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()

```

-   Germany has by far the highest trust in the legal system with a
    general upward trend
-   Responses from Poland and Slovenia fluctuate more
-   strong upward trend in Slovenia since 2014
-   lowest value in Slovenia with just above 3 in 2010
-   lowest value in Poland with just above 3 in 2020

### Unification Progress Satisfaction

#### by Country Over Time

```{r}

dt_filtered %>%
  filter(!is.na(euftf), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_euftf = mean(euftf, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_euftf, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
      scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Satisfaction with Unification by Country and Year",
    x = "Year",
    y = "Satisfatction (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()
```
    
## Ideology
### Attitudes towards Immigrants
#### by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(mean_imwbcnt = mean(imwbcnt, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_imwbcnt, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Attitudes towards Immigrants by Age Group and Country",
    x = "Age Group",
    y = "Attitudes towards Immigrants (0-10 Scale)",
    fill = "Country"
  ) +
  theme_minimal()
```

-   Polish respondents seem to be more welcoming of immigrants among all
    age groups
-   Slovenian respondents seem to be the least welcoming of immigrants
    among all age groups
-   this seems to be a contradiction to the distribution of the
    left-right scale, where a higher share of Polish respondents
    indicated to be more right, and a higher share of Slovenian
    respondents indicated to be more left
-   a boxplot in the next step could help to check for outliers
-   reasons??

#### by Country over Time --> hier auch nach Altersgruppen filtern

```{r}

dt_filtered %>%
  filter(!is.na(imwbcnt), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_imwbcnt = mean(imwbcnt, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_imwbcnt, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
      scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Immigration Attitude by Country over Time",
    x = "Year",
    y = "Immigration Attitude (0-10 Scale)",
    color = "Country"
  ) +
  theme_minimal()
```

-   there do not seem to be high outliers for positive immigration
    attitude among Polish respondents
-   Why are Polish respondents so welcoming when they are rather right?
    Would be interesting to see if the respondents who are rather right
    are also more welcoming, or if the average is high because of the
    very high immigration attitude among rather left respondents

### Trust in People
#### by Country over Time on Average

```{r}

dt_filtered %>%
  filter(!is.na(ppltrst), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_ppltrst = mean(ppltrst, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_ppltrst, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
      scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Trust in People by Country over Time",
    x = "Year",
    y = "Trust in People (0-10 Scale)",
    color = "Country"
  ) +
  theme_minimal()
```

-   lowest trust in people in Poland

### Attendance of Religious Services apart from Special Occasions

#### by Country over Time on Average

```{r}
dt_filtered %>%
  filter(!is.na(rlgatnd_rev), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_rlgatnd_rev = mean(rlgatnd_rev, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_rlgatnd_rev, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
    scale_color_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Attendance of Religious Services apart from Special Occasions",
    x = "Year",
    y = "Frequency (0–7 scale)",
    color = "Country"
  ) +
  theme_minimal()
```

-   Germany on average the country with the least attendance, followed
    by Slovenia and then Poland
-   Poland is by far the country with the highest attendance, even
    though there is an upward trend that peaked in 2020


-   Much higher share of Polish respondents attends religious services
    more often
-   Slovenia is somewhat evenly distributed
-   Germany has the highest shares of respondnets that never attend
    religious services throughout all the years
    
### Political Interest
#### by Country over Time on Average --> hier evtl. auch nach Age Groups interessant

```{r}

# Summarize mean interest per country and year
interest_trends <- dt_filtered %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%
  group_by(cntry, year) %>%
  summarise(
    polintr_rev_mean = mean(polintr_rev, na.rm = TRUE),
    .groups = "drop"
  ) 

#Reshaping to long format for plotting
trust_long <- interest_trends %>%
  pivot_longer(cols = c(polintr_rev_mean),
               names_to = "institution",
               values_to = "interest")

ggplot(interest_trends, aes(x = year, y = polintr_rev_mean, color = cntry)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Political Interest Over Time",
    x = "Year",
    y = "Political Interest (1-4 Scale)",
    color = "Country"
  ) +
  theme_minimal()

```

Interpretation of Graph on Age Groups:
-   clearly highest trust in politicians in Germany among all age groups
-   Polish respondents seem to have higher trust than Slovenian
    respondents in all age groups, except for the group 55-64

