---
title: "Project Study - Data Presemtation"
author: "Helena Zappe & Marei Göbelbecker"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-packages, eval=FALSE, include=FALSE}

# Download Packages

install.packages(c(
  "tidyverse",  
  "skimr",      
  "psych",      
  "rmarkdown",  
  "knitr",      
  "tinytex",    
  "labelled",
  "here"
  ))
tinytex::install_tinytex()  
```

```{r load_packages, include=FALSE}

# Load Packages

library(haven)
library(tidyverse)   
library(glmnet)
library(skimr)       
library(psych)       
library(rmarkdown)   
library(knitr)       
library(tinytex)     
library(labelled)
library(dplyr)
library(here)
library(tidyr)
library(purrr)

```

```{r raw_data, include=FALSE}

# 1. Lade Daten und entferne nicht benötigte Variablen
dt_all <- read_sav(here("data", "ESS_all.sav")) %>%
  select(-c(trstun))
# all binary variables: gndr, badge, bctprd, bghtprd, clsprty, contplt, dntmny, ilglpst, mmbprty, pbldmn, pbldmna, pstplonl, sgnptit, wrkorg, wrkprty, volunfp, donprty, blgetmg, rlgblg, feethngr))
```

# "What shapes trust in the European Union in Poland, Germany, and Slovenia? A comparative analysis of political attitudes, ideology, and socio-demographics."

# LASSO to find best variables

```{r}
# Step 1: Preparing Data for Lasso
# using the gilmnet package

# keep only valid observations in y
dt_filtered <- dt_all %>% drop_na(trstep)

# filter variables with too many NAs:
dt_filtered <- dt_filtered %>%
  select(where(~ !all(is.na(.)))) %>% # 100% NAs
  select(where(~ mean(is.na(.)) <= 0.8)) %>% # keep variables with less than 80% NAs
  select(-c(feethngr, loylead, lrnobed, donprty, volunfp, pbldmna, implvdm)) # these variables are not relevant enough for imputation. Also households net income has 79% NAs and is therefore not meaningful.

# checking for missing variables in percent
dt_filtered %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_pct") %>%
  filter(missing_pct > 20) %>%
  arrange(desc(missing_pct))

# defining the response Variable
y <- as.numeric(as_factor(dt_filtered$trstep))
table(y, useNA = "always")

# define the predictor variable matrix
x_data <- dt_filtered %>%
  select(-trstep) %>%
  zap_labels() %>%  # delete all labels
  mutate(across(where(is.factor), ~ as.numeric(as.character(.)))) # define all factor variables also as numeric

x_data[is.na(x_data)] <- 0  # replaces all NA in x_data

# define the matrix of response variables
x <- data.matrix(x_data)

# LASSO with Crossvalidation
lasso_cv <- cv.glmnet(x, y, alpha = 1)
```

Fitting the LASSO model:

The bigger lambda, the more coefficients will be set to 0 for variable selection.

best_lambda (lambda.min) gives the most precise model:

```{r}
# Find best Lambda
best_lambda <- lasso_cv$lambda.min

# Final LASSO Model
lasso_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)

# Lambda Precise: 0.00067 

```

lambda.1se gives the more robust, more economic (sparsamere) model:
(Oft besser generalisierbar)

```{r}
lambda_1se <- lasso_cv$lambda.1se

model_robust <- glmnet(x, y, alpha = 1, lambda = lasso_cv$lambda.1se)

# Lambda Robust: 0.00992

```

Chatty: 
"In der LASSO-Analyse wurde mit lambda.min = 0.00067 das Modell mit der höchsten Vorhersagegenauigkeit identifiziert. Dieses Modell verwendet relativ viele Prädiktoren.

Alternativ wurde mit lambda.1se = 0.00992 ein robusteres Modell berechnet, das sparsamer ist und potenziell besser generalisiert. Hier bleiben nur die wichtigsten Prädiktoren erhalten.

Je nach Forschungsziel (Präzision vs. Interpretierbarkeit) kann eines der beiden Modelle bevorzugt werden."







