---
title: "Project Study"
author: "Marei Göbelbecker & Helena Zappe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-packages, eval=FALSE, include=FALSE}

# Download Packages

install.packages(c(
  "tidyverse",  
  "skimr",      
  "psych",      
  "rmarkdown",  
  "knitr",      
  "labelled",
  "here"
  ))
```

```{r load_packages, include=FALSE}

# Load Packages

library(haven)
library(here)
library(tidyverse)   
library(skimr)       
#library(psych)       
library(rmarkdown)  
library(knitr)       
library(labelled)
library(dplyr)
library(tidyr)
#library(forcats)
library(glmnet)
```

```{r raw_data, include=FALSE}

# Load Raw Data

dt_raw <- read_sav(here("data","ESS_all.sav"))

```

`"What factors shapes trust in the European Union in Poland, Germany, and Slovenia? A comparative analysis of political attitudes, ideology, and socio-demographics."`

*Note: We use the terms "Trust in the EU" and "Trust in the European Parliament" synonymously.*

# Data Wrangling

In order to work with our dataset, we need to reshape some variables. We are filtering for our three countries, recoding the age variable into age groups, reversing the scale for political interest (such that a high number on the scale means one has high interest in politics) and are winsurizing the education years to get rid of outliers.

```{r data_preparation}

dt_all <- dt_raw %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%  
  mutate(
    age_group = case_when(
      agea < 18 ~ "<18",                      
      agea >=18 & agea <= 24 ~ "18-24",
      agea > 24 & agea <= 34 ~ "25-34",
      agea > 34 & agea <= 44 ~ "35-44",
      agea > 44 & agea <= 54 ~ "45–54",
      agea > 54 & agea <= 64 ~ "55-64",
      agea > 64 ~ "65+"
    ))  %>%                                   
  mutate(
    age_group = factor(age_group, levels = c("<18", "18-24", "25-34", "35-44", "45–54", "55-64", "65+")),
    polintr_rev = 5 - polintr  # higher = more interested               
    ) %>%
  mutate(year = case_when(    # Mapping ESS round to the year they were conducted in in "year"
    essround == 1 ~ 2002,
    essround == 2 ~ 2004,
    essround == 3 ~ 2006,
    essround == 4 ~ 2008,
    essround == 5 ~ 2010,
    essround == 6 ~ 2012,
    essround == 7 ~ 2014,
    essround == 8 ~ 2016,
    essround == 9 ~ 2018,
    essround == 10 ~ 2020,
    essround == 11 ~ 2022,
    TRUE ~ NA_real_
    )) %>%
    select(-c(agea, polintr, essround, trstun, trstplc, trstplt, trstprl, trstprt, idno, proddate, bctprd, wrkorg, edition, name, pplfair, pplhlp, stfdem, imueclt, imdfetn, impcntr, imbgeco, badge)) # variables that are not relevant for interpretation


# winsurize education years to get rid of outliers

# Calculate thresholds
q_lower <- quantile(dt_all$eduyrs, 0.01, na.rm = TRUE)
q_upper <- quantile(dt_all$eduyrs, 0.99, na.rm = TRUE)

# Winsorize: replace values outside bounds
dt_all <- dt_all %>%
  mutate(eduyrs_winsor = case_when(
    eduyrs < q_lower ~ q_lower,
    eduyrs > q_upper ~ q_upper,
    TRUE ~ eduyrs
  )) %>%
  select(-eduyrs)
```

```{r new_dataset, eval=FALSE}
# Saving Data as new Dataset for Lasso
write_sav(dt_all, here("data", "ESS_prepared.sav"))
```

# Run LASSO to find the Best Variables for Analysis

```{r prepare_lasso}
# Step 1: Preparing Data for Lasso
# using the gilmnet package

# filter variables with too many NAs:
dt_filtered <- dt_all %>%
  drop_na(trstep) %>% # keep only valid observations in y
  select(where(~ !all(is.na(.)))) %>% # delete where is 100% NAs
  select(where(~ mean(is.na(.)) <= 0.5)) %>% # only keep variables with less than 50% NAs
  
# checking for missing variables in percent
dt_filtered %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_pct") %>%
  filter(missing_pct > 20) %>%
  arrange(desc(missing_pct))

# accounting for differences in country size and selection
dt_filtered <- dt_filtered %>% drop_na(anweight) # dropping all obs where anweight is NA
weights <- dt_filtered$anweight # establishing the weighting

# defining the response Variable
y <- as.numeric(as_factor(dt_filtered$trstep)) # vector saved as numeric
table(y, useNA = "always")

# define the predictor variable matrix
x_data <- dt_filtered %>%
  select(-c(trstep, anweight, pspwght, dweight, pweight )) %>% 
  zap_labels() %>%  # delete all labels
  mutate(across(where(is.factor), ~ as.numeric(as.character(.)))) # define all factor variables also as numeric

x_data[is.na(x_data)] <- 0  # impute / replace all NA with zero

# define the matrix of response variables
x <- data.matrix(x_data)

# Set seed for reproducibility
set.seed(123)

# LASSO with Crossvalidation and weighting
lasso_cv <- cv.glmnet(x, y, alpha = 1, weights = weights)

```

## Fitting the LASSO model

Now we need to find the best lasso model for fit and interpretability. We want a robust, not overfitted model which can still be easily interpreted with its amount of variables.

Before fitting the lasso model, we would suggest that the variables shaping trust in the EU would be: ppltrst, euftf, freehms, stfedu, stfgov, stfhlth, trstlgl, vote, imwbcnt, health, rlgatnd, gndr, year

To see if we are correct, we first need to find the lambda penalty term, that is used for variable selection. For further understanding it is important to know, that the bigger lambda gets, the more coefficients will be set to zero for the variable selection. Therefore we fitted the "best lambda model" with the minimal lambda and MSE, which gives the most precise model. With a Lambda of 0.0039 there are still 73 varaibles in the model, which is not well interpretable and probably overfitted. In addition, we ran a more robust and economic model by using lambda1.se which makes it easier to generalize. In this model, lambda is set higher at 0.0280 and also allows the MSE to be higher, which leaves us with 37 variables. This is already a lot better to interpret, but would still exceed the frame of this project. Therefore, in the last step, we fixed the number of variables and looked, whether lambda is still small enough to be used for interpretation. We decided to stay on le lower bound of 10 to 30 variables and tried several constellations. Ran over which variables are collinear or not useful for interpretation, to delete them before running lasso again. With this we came to the conclusion, that 13 variables with a lambda of 0.0821 will give us the best model for this project.

```{r lasso_model_max}

## Fixing Number of variables to get an interpretable model

lasso_model_max <- glmnet(x, y, alpha = 1, dfmax = 14) 

# Extracting the Coefficients
coef_lasso <- coef(lasso_model_max)

# Anzahl ≠ 0-Koeffizienten je Lambda (inkl. Intercept)
nonzero <- lasso_model_max$df
lambda_index <- which(nonzero == 14)[1]  # 13 Predictors + 1 Intercept

# Wenn gefunden:
if (!is.na(lambda_index)) {
  lambda_max <- lasso_model_max$lambda[lambda_index]
  coef_max <- coef(lasso_model_max, s = lambda_max)
  vars_max <- rownames(coef_max)[which(coef_max[, 1] != 0)]
  vars_max <- vars_max[vars_max != "(Intercept)"]
  print(vars_max)
} else {
  message("No Lambda found.")
}
```

```{r new_dataset, eval=FALSE}
### Saving the new subset with original labels and coding

# Creating a new dataset with original structure but only our selected predictors and the outcome variable
  dt_lasso <- dt_all %>%
    select(all_of(vars_max), trstep)  

# Keeping only valid cases of trstep
  dt_lasso <- dt_lasso %>%
    drop_na(trstep)  
  
# Saving it 
write_sav(dt_lasso, here("data", "ESS_lasso.sav"))
```

The variables we will work with are: **cntry** (country), **ppltrst** (Most people can be trusted or you can't be too careful), **clsprty** (Feel closer to a particular party than all others), **euftf** (European Union: European unification go further or gone too far), **freehms** (Gays and lesbians free to live as they wish), **stfedu** (state of education in country nowadays), **stfgov** (How satisfied with the national government), **stfhlth** (State of health services in country nowadays), **trstlgl** (Trust in the legal system), **imwbcnt** (Immigrants make country worse or better place to live), **rlgatnd** (How often attend religious services apart from special occations), **gndr** (Gender), **age_group** (Age group), **year** (Ess round)

Our hypothesis: Dependend on country, Gender and Age and differing by Year... ...the more politically interested and the more left on the political scale one is, the more they trust the European Parliament. ...the more open towards other people one is, the more one trusts the EU. ...the more satisfied one is with the state of their own county nowadays, the more one trusts in the EP.

-   Political attitudes: clsprty, euftf, stfgov, trstlgl
-   Ideology: ppltrst, freehms, imwbcnt, rlgatnd
-   Socio-demographics: stfedu, stfhlth, gndr, age_group
-   Other: cntry, year

# Descriptive Statistics

## Summary Statistics of Variables by Country

```{r}
dt_lasso_labeled %>%
  group_by(cntry) %>%
  summarise(across(
    c(clsprty, euftf, stfgov, trstlgl, ppltrst, freehms, imwbcnt, rlgatnd, stfedu, stfhlth, gndr, age_group, year, trstep),
    list(mean = ~mean(.x, na.rm = TRUE))#,
         #median = ~median(.x, na.rm = TRUE),
         #Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         #Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  )) 

```

- People feel closest to a particular party in Poland and close behind in Slovenia, Slightly less than half of the people feel close in Germany.
- People in Poland think, that Unification should go further, in the other two countries as well, but not as much.
- People in Germany are most satisfied with their national government and least satisfied in Poland.
- People in Germany trust their legal system way more than the other two, the others are very close together.
- People in Poland are least trusting towards others and most trusting in Germany.
- People in Germany are the most open towards homosexuality, while the others are close together. 
- Polish people are least positive towards immigrants, while people in Slovenia are a little more positive.
- Polish people are by far least religious, and most in Germany.
- State of education seems to be worst in Germany and best in Slovenia.
- Health system seems to be worst in Poland and best in Germany.

evtl. freehms scale umdrehen


## Summary Statistics of Trust in EU by Country

```{r}
dt_lasso_labeled %>%
  group_by(cntry) %>%
  summarise(across(
    c(trstep),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  ))

```

-   Mean trust in EU is rather similar across all three countries

## Visualization

### Trust in EU Time for all three Countries 

```{r}

ggplot(trust_trends, aes(x = year, y = trust_EU, color = cntry)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "DE" = "gold",  # Germany
      "PL" = "red",  # Poland
      "SI" = "blue"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Trust in the EU Over Time",
    x = "Year",
    y = "Mean Trust in EU",
    color = "Country"
  ) +
  theme_minimal()


```

-   Strong fluctuations over time in all three countries.
-   Mean trust in Slovenia peaked in 2006 and was the lowest in 2014.
-   Mean trust in Poland peaked in 2002 and was the lowest in 2014.
-   Mean trust in Germany peaked in 2023 and was the lowest in 2010.
-   Slovenia shows the highest fluctuations in mean trust in the EU, with a sharp decline between 2006 and 2014.

### Distribution of Age Groups - Ab hier weiter bearbeiten!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}

dt_filtered %>%
  count(cntry, age_group) %>%
  group_by(cntry) %>%
  mutate(prop = n / sum(n) * 100) %>%
  ggplot(aes(x = cntry, y = prop, fill = age_group)) +
  geom_col(position = "dodge") +
  labs(
    title = "Relative Distribution of Age Groups by Country",
    x = "Country",
    y = "Proportion of Respondents",
    fill = "Age Group"
  ) +
  theme_minimal()

```

-   Among all countries, the age group "65+" has the highest share
-   Age distribution has an upward trend in all three countries, indicating an aging population

### Average Trust in EU by Age Group and Country

```{r}
dt_filtered %>%
  group_by(cntry, age_group) %>%
  summarise(mean_trustEU = mean(trstep, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_trustEU, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",     # Germany - Blue
      "PL" = "#f6ae2d",     # Poland - Light Orange
      "SI" = "#f26419"      # Slovenia - Dark Orange/Pink
    ),
    labels = c(
      "DE" = "Germany",
      "PL" = "Poland",
      "SI" = "Slovenia"
    )
  ) +
  labs(
    title = "Average Trust in EU by Age Group and Country",
    x = "Age Group",
    y = "Mean Trust in EU",
    fill = "Country"
  ) +
  theme_minimal()

```

-   highest mean trust in EU among people younger than 18 in all three countries, with respondents from Germany having the highest trust
-   general downward trend in meant trust in EU across all countries up to the age group 55-64
-   trust increases again for respondents being 65 or older

### Average Trust in Politicians by Age Group and Country

```{r}
dt_filtered %>%
  group_by(cntry, age_group) %>%
  summarise(mean_trustpol = mean(polintr_rev, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_trustpol, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",     # Germany - Blue
      "PL" = "#f6ae2d",     # Poland - Light Orange
      "SI" = "#f26419"      # Slovenia - Dark Orange/Pink
    ),
    labels = c(
      "DE" = "Germany",
      "PL" = "Poland",
      "SI" = "Slovenia"
    )
  ) +
  labs(
    title = "Average Trust in Politicians by Age Group and Country",
    x = "Age Group",
    y = "Mean Trust in Politicians",
    fill = "Country"
  ) +
  theme_minimal()
```

-   clearly highest trust in politicians in Germany among all age groups
-   Polish respondents seem to have higher trust than Slovenian respondents in all age groups, except for the group 55-64 

### Left-Right Political Placement across Countries

```{r}

# Filtering to valid left-right scale values
lr_data <- dt_filtered %>%
  filter(!is.na(lrscale)) %>%
  group_by(cntry, lrscale) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(lr_data, aes(x = cntry, y = percentage, fill = factor(lrscale, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "Left-Right Scale"
  ) +
  labs(
    title = "Placement on Left-Right Political Scale",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

-   highest share of respondents selected 5 as their political attitude
-   higher share of respondents in Germany and Slovenia are rather left
-   higher share of respondents in Poland are rather right

### Satisfaction with Government by Country

```{r}

ggplot(dt_filtered, aes(x = cntry, y = stfgov, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Satisfaction with Government by Country",
    y = "Satisfaction with Government (Higher = More)",
    x = "Country",
    fill = "Country"
  ) +
  theme_minimal()


```

-   satisfation with democracy highest in Germany and similar in Slovenia and Poland

### Education Years per Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = eduyrs_winsor, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Education Years by Country",
       y = "Years of Education", x = "Country")

```

- we winsorized outliers earlier
- highest years of education in Germany

### Years of Education by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(mean_eduyrs = mean(eduyrs_winsor, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_eduyrs, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Mean Years of Education by Age Group and Country",
    x = "Age Group",
    y = "Mean Years of Education",
    fill = "Country"
  ) +
  theme_minimal()
```

-   graph illustrates differences among age groups across countries
-   Germany has the highest average years of education among all age groups, followed by Poland

### Attitudes towards Immigrants by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(mean_imwbcnt = mean(imwbcnt, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_imwbcnt, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Attitudes towards Immigrants by Age Group and Country",
    x = "Age Group",
    y = "Attitudes towards Immigrants",
    fill = "Country"
  ) +
  theme_minimal()
```

-   Polish respondents seem to be more welcoming of immigrants among all age groups
-   Slovenian respondents seem to be the least welcoming of immigrants among all age groups
-   this seems to be a contradiction to the distribution of the left-right scale, where a higher share of Polish respondents indicated to be more right, and a higher share of Slovenian respondents indicated to be more left
-   a boxplot in the next step could help to check for outliers
-   reasons??

### Immigration Attitude per Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = imwbcnt, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Immigration Attitude by Country",
       y = "Immigration Attitude", x = "Country")
```

-   there do not seem to be high outliers for positive immigration attitude among Polish respondents
-   Why are Polish respondents so welcoming when they are rather right? Would be interesting to see if the respondents who are rather right are also more welcoming, or if the average is high because of the very high immigration attitude among rather left respondents




