---
title: "Project Study"
author: "Marei Göbelbecker & Helena Zappe"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-packages, eval=FALSE, include=FALSE}

# Download Packages

install.packages(c(
  "tidyverse",  
  "skimr",      
  "psych",      
  "rmarkdown",  
  "knitr",      
  "labelled",
  "here"
  ))
```

```{r load_packages, include=FALSE}

# Load Packages

library(haven)
library(here)
library(tidyverse)   
library(skimr)       
#library(psych)       
library(rmarkdown)  
library(knitr)       
library(labelled)
library(dplyr)
library(tidyr)
#library(forcats)
library(glmnet)
library(ggplot2)

```

```{r raw_data, include=FALSE}

# Load Raw Data

dt_raw <- read_sav(here("data","ESS_all.sav"))

```

`"What factors shapes trust in the European Union in Poland, Germany, and Slovenia? A comparative analysis of political attitudes, ideology, and socio-demographics."`

*Note: We use the terms "Trust in the EU" and "Trust in the European
Parliament" synonymously.*

# Data Wrangling

In order to work with our dataset, we need to reshape some variables. We
are filtering for our three countries, recoding the age variable into
age groups, reversing the scale for political interest (such that a high
number on the scale means one has high interest in politics) and are
winsurizing the education years to get rid of outliers.

```{r data_preparation}
dt_all <- dt_raw %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%  
  mutate(
    age_group = case_when(
      agea < 18 ~ "<18",                      
      agea >=18 & agea <= 24 ~ "18-24",
      agea > 24 & agea <= 34 ~ "25-34",
      agea > 34 & agea <= 44 ~ "35-44",
      agea > 44 & agea <= 54 ~ "45–54",
      agea > 54 & agea <= 64 ~ "55-64",
      agea > 64 ~ "65+"
    )) %>%
  mutate(
    age_group = factor(age_group, levels = c("<18", "18-24", "25-34", "35-44", "45–54", "55-64", "65+"))
    ) %>%
  mutate(
    polintr_rev = 5 - polintr  # higher = more interested  
    ) %>%
  mutate(
    freehms_rev = 6 - freehms  # higher = more agreed
    ) %>%
  mutate(
    rlgatnd_rev = 8- rlgatnd   # higher = more often
  ) %>%
  mutate(
    health_rev = 6 - health    # higher = better
  ) %>%
  mutate(year = case_when(    # Mapping ESS round to the year they were conducted in in "year"
    essround == 1 ~ 2002,
    essround == 2 ~ 2004,
    essround == 3 ~ 2006,
    essround == 4 ~ 2008,
    essround == 5 ~ 2010,
    essround == 6 ~ 2012,
    essround == 7 ~ 2014,
    essround == 8 ~ 2016,
    essround == 9 ~ 2018,
    essround == 10 ~ 2020,
    essround == 11 ~ 2022,
    TRUE ~ NA_real_
    )) %>%
    select(-c(polintr, freehms, , health, rlgatnd, essround, trstun, trstplc, trstplt, trstprl, trstprt, idno, proddate, bctprd, wrkorg, edition, name, pplfair, pplhlp, stfdem, imueclt, imdfetn, impcntr, imbgeco, badge)) # variables that are not relevant for interpretation 


# winsurize education years to get rid of outliers

# Calculate thresholds
q_lower <- quantile(dt_all$eduyrs, 0.01, na.rm = TRUE)
q_upper <- quantile(dt_all$eduyrs, 0.99, na.rm = TRUE)

# Winsorize: replace values outside bounds
dt_all <- dt_all %>%
  mutate(eduyrs_winsor = case_when(
    eduyrs < q_lower ~ q_lower,
    eduyrs > q_upper ~ q_upper,
    TRUE ~ eduyrs
  )) %>%
  select(-eduyrs)
```

```{r new_dataset, eval=FALSE}
# Saving Data as new Dataset for Lasso
write_sav(dt_all, here("data", "ESS_prepared.sav"))
```

# Run LASSO to find the Best Variables for Analysis

```{r prepare_lasso}
# Step 1: Preparing Data for Lasso
# using the gilmnet package

# filter variables with too many NAs:
dt_filtered <- dt_all %>%
  drop_na(trstep) %>% # keep only valid observations in y
  select(where(~ !all(is.na(.)))) %>% # delete where is 100% NAs
  mutate(across(everything(), ~ ifelse(. %in% c(77,88,99), NA, .))) %>% # refusals and no anser which are not coded as NAs
  select(where(~ mean(is.na(.)) <= 0.5)) # only keep variables with less than 50% NAs
  
# checking for missing variables in percent
dt_filtered %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_pct") %>%
  filter(missing_pct > 20) %>%
  arrange(desc(missing_pct))

# accounting for differences in country size and selection
dt_filtered <- dt_filtered %>% drop_na(anweight) # dropping all obs where anweight is NA
weights <- dt_filtered$anweight # establishing the weighting

# defining the response Variable
y <- as.numeric(as_factor(dt_filtered$trstep)) # vector saved as numeric
table(y, useNA = "always")

# define the predictor variable matrix
x_data <- dt_filtered %>%
  select(-c(trstep, anweight, pspwght, dweight, pweight, age_group)) %>% # age_group not interpretable by lasso, therefore keep agea but later use age_group for interpretation
  zap_labels() %>%  # delete all labels
  mutate(across(where(is.factor), ~ as.numeric(as.character(.)))) # define all factor variables also as numeric

x_data[is.na(x_data)] <- 0  # impute / replace all NA with zero

# define the matrix of response variables
x <- data.matrix(x_data)

# Set seed for reproducibility
set.seed(123)

# LASSO with Crossvalidation and weighting
lasso_cv <- cv.glmnet(x, y, alpha = 1, weights = weights)

```

## Fitting the LASSO model

Now we need to find the best lasso model for fit and interpretability.
We want a robust, not overfitted model which can still be easily
interpreted with its amount of variables.

Before fitting the lasso model, we would suggest that the variables
shaping trust in the EU would be: ppltrst, euftf, freehms, stfedu,
stfgov, stfhlth, trstlgl, vote, imwbcnt, health, rlgatnd, gndr, year

To see if we are correct, we first need to find the lambda penalty term,
that is used for variable selection. For further understanding it is
important to know, that the bigger lambda gets, the more coefficients
will be set to zero for the variable selection. Therefore we fitted the
"best lambda model" with the minimal lambda and MSE, which gives the
most precise model. With a Lambda of 0.0039 there are still 73 varaibles
in the model, which is not well interpretable and probably overfitted.
In addition, we ran a more robust and economic model by using lambda1.se
which makes it easier to generalize. In this model, lambda is set higher
at 0.0280 and also allows the MSE to be higher, which leaves us with 37
variables. This is already a lot better to interpret, but would still
exceed the frame of this project. Therefore, in the last step, we fixed
the number of variables and looked, whether lambda is still small enough
to be used for interpretation. We decided to stay on le lower bound of
10 to 30 variables and tried several constellations. Ran over which
variables are collinear or not useful for interpretation, to delete them
before running lasso again. With this we came to the conclusion, that 11
variables with a lambda of 0.0901 will give us the best model for this
project.

```{r lasso_model_max}
set.seed(123)

## Fixing Number of variables to get an interpretable model

lasso_model_max <- glmnet(x, y, alpha = 1, dfmax = 12) 

# Extracting the Coefficients
coef_lasso <- coef(lasso_model_max)

# Anzahl ≠ 0-Koeffizienten je Lambda (inkl. Intercept)
nonzero <- lasso_model_max$df
lambda_index <- which(nonzero == 12)[1]  # 11 Predictors + 1 Intercept

# Wenn gefunden:
if (!is.na(lambda_index)) {
  lambda_max <- lasso_model_max$lambda[lambda_index]
  coef_max <- coef(lasso_model_max, s = lambda_max)
  vars_max <- rownames(coef_max)[which(coef_max[, 1] != 0)]
  vars_max <- vars_max[vars_max != "(Intercept)"]
  print(vars_max)
} else {
  message("No Lambda found.")
}
```
 
```{r new_dataset, eval=FALSE}
### Saving the new subset with original labels and coding

# Creating a new dataset with original structure but only our selected predictors and the outcome variable
  dt_lasso <- dt_all %>%
    select(all_of(vars_max), trstep)  

# Keeping only valid cases of trstep
  dt_lasso <- dt_lasso %>%
    drop_na(trstep)  
  
# Saving it 
write_sav(dt_lasso, here("data", "ESS_lasso.sav"))
```

The variables we will work with are: **cntry** (country), **ppltrst**
(Most people can be trusted or you can't be too careful) , **euftf** (European
Union: European unification go further or gone too far) , **stfedu** (state of
education in country nowadays), **stfgov** (How satisfied with the
national government), **stfhlth** (State of health services in country
nowadays), **trstlgl** (Trust in the legal system), **imwbcnt**
(Immigrants make country worse or better place to live), **rlgatnd_rev**
(How often attend religious services apart from special occations),
**gndr** (Gender), **age_group** (Age group), **year** (Ess round)


*Nochmal überarbeiten*
Our hypothesis: Dependend on country, Gender and Age and differing by
Year... 

 ...the more satisfied one is with the state of their own county nowadays, the more one trusts in the EP.

...the more open towards other people one is, the more one trusts the EU.

...people who want the European unification to go even further, also trust more in the EU.


-   Political attitudes: stfedu, stfhlth, euftf, stfgov, trstlgl
-   Ideology: ppltrst, imwbcnt, rlgatnd_rev
-   Socio-demographics: gndr, age_group
-   Other: cntry, year



# Descriptive Statistics

## Summary Statistics of Variables by Country

```{r}
summary_table <- dt_filtered %>%
  group_by(cntry) %>%
  summarise(across(
    c(trstep, ppltrst, euftf, stfedu, stfgov, stfhlth, trstlgl, imwbcnt, rlgatnd_rev, gndr, agea), # here use agea, but later always age_group
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  )) 

View(as.data.frame(summary_table))

```

Analysis on average across one country and all years.
-   Trust in the EU is very similar.
-   People in Poland are least trusting towards others and most trusting
    in Germany.
-   People in Poland think, that Unification should go further, in the
    other two countries as well, but not as much. While all are very similarly rather indifferent.
-   The state of Education seems to be best in Slovenia, and worst in Germany.
-   People in Germany are most satisfied with their national government
    and least satisfied in Slovenia.
-   People in Germany trust their legal system way more than the other
    two, the others are very close together.
-   The Polish health system seems to be the worst and best in Germany.
-   Germans trust their legal system far more than the other two countries, which seem to be very similarly untrusting.
-   Polish people are most and slovenians are the least positive towards immigrants.
-   Polish people are by far most religious, and least in Germany.

## Summary Statistics of Trust in EU by Country

```{r}
dt_filtered %>%
  group_by(cntry) %>%
  summarise(across(
    c(trstep),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  ))

```

-   Mean trust in EU is rather similar across all three countries

# Visualization - Other interesting variables

## Distribution of Voting Behavior - Not necessary, since not one of our predictors

```{r}

# Filtering to valid left-right scale values
vote_data <- dt_filtered %>%
  filter(!is.na(vote)) %>%
  group_by(cntry, vote) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(vote_data, aes(x = cntry, y = percentage, fill = factor(vote, levels = rev(0:3)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "1" = "#002060", "2" = "#4A90E2", "3" = "#DDDDDD"
    ),
    name = "Voted Last National Election"
  ) +
  labs(
    title = "Distribution of Voting Behavior",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

-   highest proportion of voters in Germany
-   Slovenia has the highest share of people that did not vote

## Left-Right Political Placement across Countries - Also not one of our predictors

```{r}

# Filtering to valid left-right scale values
lr_data <- dt_filtered %>%
  filter(!is.na(lrscale)) %>%
  group_by(cntry, lrscale) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(lr_data, aes(x = cntry, y = percentage, fill = factor(lrscale, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "Left-Right Scale"
  ) +
  labs(
    title = "Placement on Left-Right Political Scale",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

-   highest share of respondents selected 5 as their political attitude
-   higher share of respondents in Germany and Slovenia are rather left
-   higher share of respondents in Poland are rather right

## Years of Education

### by Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = eduyrs_winsor, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Education Years by Country",
       y = "Years of Education", x = "Country")

```

-   we winsorized outliers earlier
-   highest years of education in Germany

### by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(mean_eduyrs = mean(eduyrs_winsor, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_eduyrs, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Mean Years of Education by Age Group and Country",
    x = "Age Group",
    y = "Mean Years of Education",
    fill = "Country"
  ) +
  theme_minimal()
```

-   graph illustrates differences among age groups across countries
-   Germany has the highest average years of education among all age
    groups, followed by Poland

# Visualization - Dependent Variable

## Trust in EU  

### by Country over Time

```{r}
# Summarize mean trust per country and year
trust_trends <- dt_filtered %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%
  group_by(cntry, year) %>%
  summarise(
    trust_EU = mean(trstep, na.rm = TRUE),
    .groups = "drop"
  ) 

#Reshaping to long format for plotting
trust_long <- trust_trends %>%
  pivot_longer(cols = c(trust_EU),
               names_to = "institution",
               values_to = "trust")
#Plot
ggplot(trust_trends, aes(x = year, y = trust_EU, color = cntry)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "DE" = "#33658a",  # Germany
      "PL" = "#f6ae2d",  # Poland
      "SI" = "#f26419"   # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Trust in the EU Over Time",
    x = "Year",
    y = "Mean Trust in EU",
    color = "Country"
  ) +
  theme_minimal()


```

-   strong flucutations over time fro all three countries
-   mean trust in Slovenia peaked in 2006 and was the lowest in 2014
-   mean trust in Poland peaked in 2002 and was the lowest in 2014
-   mean trust in Germany peaked in 2023 and was the lowest in 2010
-   Slovenia shows the highest fluctuations in mean trust in the EU,
    with a sharp decline between 2006 and 2014
    
### by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(trstep = mean(trstep, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = trstep, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Average Trust in the EU by Age Group and Country",
    x = "Age Group",
    y = "Mean Trust in the EU",
    fill = "Country"
  ) +
  theme_minimal()
```

-   highest mean trust in EU among people younger than 18 in all three
    countries, with respondents from Germany having the highest trust
-   general downward trend in meant trust in EU across all countries up
    to the age group 55-64
-   trust increases again for respondents being 65 or older

# Visualization - Sociodemographic Predictors

## Distribution of Gender by Country

```{r}

dt_filtered %>%
  mutate(gndr = as_factor(gndr)) %>%  # convert haven_labelled to factor
  count(cntry, gndr) %>%
  group_by(cntry) %>%
  mutate(prop = n / sum(n) * 100) %>%
  ggplot(aes(x = cntry, y = prop, fill = gndr)) +
  geom_col(position = "dodge") +
  labs(
    title = "Relative Distribution of Gender by Country",
    x = "Country",
    y = "Proportion of Respondents",
    fill = "Gender"
  ) +
  theme_minimal()

```

-   distribution among all three countries rather similar around 50% for
    each gender
-   Germany has slightly more male respondents
-   Poland and Slovenia have slightly more female respondents
-   gap is the highest in Slovenia

## Distribution of Age Groups - Tabelle reicht eigentlich, da Graph nicht aussagekräftig

```{r}
table(dt_filtered$age_group)

dt_filtered %>% 
  count(cntry, age_group) %>% 
  group_by(cntry) %>% 
  mutate(prop = n / sum(n) * 100) %>% 
  ggplot(aes(x = age_group, y = prop, fill = cntry)) +
  geom_col(position = "dodge") +
  labs(
    title = "Age Group Distribution Within Each Country",
    x = "Age Group",
    y = "Proportion of Respondents (%)",
    fill = "Country"
  ) +
  theme_minimal()

```

-   Among all countries, the age group "65+" has the highest share
-   Age distribution has an upward trend in all three countries,
    indicating an aging population


# Visualization - Political Attitudes

## Satisfaction with National Government

### by Country Over Time

```{r}

# Ensure year is treated properly
dt_filtered %>%
  filter(!is.na(stfgov), !is.na(year)) %>%
  group_by(cntry, year, stfgov) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry, year) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = cntry, y = percentage, fill = factor(stfgov, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  facet_wrap(~year) +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "Satisfaction"
  ) +
  labs(
    title = "Satisfaction with National Government by Country over Time",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

-   satisfation with national government fluctuated over time among all
    countries
-   Polish respondents were very dissatisfied in 2020
-   German respondents were very dissatisfied in 2002
-   Slovenian respondents were very dissatisfied in 2012

### by Country over Time on Average

```{r}
dt_filtered %>%
  filter(!is.na(stfgov), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_stfgov = mean(stfgov, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_stfgov, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Average Satisfaction with National Government Over Time",
    x = "Year",
    y = "Average Satisfaction (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()

```

## Satisfaction with State of Education by Country

### by Country On Average

```{r}
# Filtering to valid scale values
edu_data <- dt_filtered %>%
  filter(!is.na(stfedu)) %>%
  group_by(cntry, stfedu) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(edu_data, aes(x = cntry, y = percentage, fill = factor(stfedu, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "State of Education in Country Nowadays"
  ) +
  labs(
    title = "Satisfaction with State of Education by Country",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

-   highest satisfaction with education to be seen in Slovenia, followed
    by Poland and then Germany

### by Country Over Time

```{r}
# Ensure year is treated properly
dt_filtered %>%
  filter(!is.na(stfedu), !is.na(year)) %>%
  group_by(cntry, year, stfedu) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry, year) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = cntry, y = percentage, fill = factor(stfedu, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  facet_wrap(~year) +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "Satisfaction"
  ) +
  labs(
    title = "Satisfaction with State of Education by Country and Year",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

## Satisfaction with the State of Health within Country

### by Country over Time

```{r}

# Ensure year is treated properly
dt_filtered %>%
  filter(!is.na(stfhlth), !is.na(year)) %>%
  group_by(cntry, year, stfhlth) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry, year) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = cntry, y = percentage, fill = factor(stfhlth, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  facet_wrap(~year) +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "Satisfaction"
  ) +
  labs(
    title = "Satisfaction with State of Health by Country over Time",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### by Country over Time on Average

```{r}
dt_filtered %>%
  filter(!is.na(stfhlth), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_stfhlth = mean(stfhlth, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_stfhlth, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Average Satisfaction with State of Health by Country over Time",
    x = "Year",
    y = "Average Satisfaction (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()
```

-   Germany peaked around 2016 just above 6, also the highest value
    among all three countries, but satisfaction declined after that
-   Polish satisfaction crashed to under 3 in 2020
-   values from Polish respondents are the lowest among all three
    countries

## Trust in the Legal System by Country over Time on Average

```{r}

dt_filtered %>%
  filter(!is.na(trstlgl), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_trstlgl = mean(trstlgl, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_trstlgl, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Trust in the legal System by Country over Time",
    x = "Year",
    y = "Average State (0–10 scale)",
    color = "Country"
  ) +
  theme_minimal()

```

-   Germany has by far the highest trust in the legal system with a
    general upward trend
-   Responses from Poland and Slovenia fluctuate more
-   strong upward trend in Slovenia since 2014
-   lowest value in Slovenia with just above 3 in 2010
-   lowest value in Poland with just above 3 in 2020


# Visualization - Ideology

## Attitudes towards Immigrants

### by Age Group and Country

```{r}
dt_filtered %>%
  group_by(age_group, cntry) %>%
  summarise(mean_imwbcnt = mean(imwbcnt, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = age_group, y = mean_imwbcnt, fill = cntry)) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany
      "PL" = "#f6ae2d",   # Poland
      "SI" = "#f26419"    # Slovenia
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(
    title = "Attitudes towards Immigrants by Age Group and Country",
    x = "Age Group",
    y = "Attitudes towards Immigrants",
    fill = "Country"
  ) +
  theme_minimal()
```

-   Polish respondents seem to be more welcoming of immigrants among all
    age groups
-   Slovenian respondents seem to be the least welcoming of immigrants
    among all age groups
-   this seems to be a contradiction to the distribution of the
    left-right scale, where a higher share of Polish respondents
    indicated to be more right, and a higher share of Slovenian
    respondents indicated to be more left
-   a boxplot in the next step could help to check for outliers
-   reasons??

### per Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = imwbcnt, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Immigration Attitude by Country",
       y = "Immigration Attitude", x = "Country")
```

-   there do not seem to be high outliers for positive immigration
    attitude among Polish respondents
-   Why are Polish respondents so welcoming when they are rather right?
    Would be interesting to see if the respondents who are rather right
    are also more welcoming, or if the average is high because of the
    very high immigration attitude among rather left respondents

## Trust in People by Country

```{r}
ggplot(dt_filtered, aes(x = cntry, y = ppltrst, fill = cntry)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "DE" = "#33658a",   # Germany - Blue
      "PL" = "#f6ae2d",   # Poland - Light Orange
      "SI" = "#f26419"    # Slovenia - Dark Orange/Pink
    ),
    labels = c("DE" = "Germany", "PL" = "Poland", "SI" = "Slovenia")
  ) +
  labs(title = "Trust in People by Country",
       y = "Trust", x = "Country")
```

-   lowest trust in people in Poland

## Unification Progress Satisfaction

### by Country on Average

Higher values indicate a stronger desire for unification.

```{r}
# Filtering to valid scale values
eu_data <- dt_filtered %>%
  filter(!is.na(euftf)) %>%
  group_by(cntry, euftf) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry) %>%
  mutate(percentage = count / sum(count) * 100)


#Plotting the diverging stacked bar chart
ggplot(eu_data, aes(x = cntry, y = percentage, fill = factor(euftf, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "European Unification Go Further or Gone Too Far"
  ) +
  labs(
    title = "Satisfaction with Unification by Country",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

-   in all three countries, the majority of repsondents voted values
    above average, indicating they want the unification to go further

### by Country Over Time

```{r}
# Ensure year is treated properly
dt_filtered %>%
  filter(!is.na(euftf), !is.na(year)) %>%
  group_by(cntry, year, euftf) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry, year) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = cntry, y = percentage, fill = factor(euftf, levels = rev(0:10)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  facet_wrap(~year) +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C", "8" = "#FF914D", "9" = "#C23B22", "10" = "#800000"
    ),
    name = "European Unification Go Further or Gone Too Far"
  ) +
  labs(
    title = "Satisfaction with Unification by Country and Year",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```


## Attendance of Religious Services apart from Special Occasions

### by Country over Time

Lower values indicate that a respondent attends more often (reverse
scale?)

```{r}
# Ensure year is treated properly
dt_filtered %>%
  filter(!is.na(rlgatnd_rev), !is.na(year)) %>%
  group_by(cntry, year, rlgatnd_rev) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(cntry, year) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ggplot(aes(x = cntry, y = percentage, fill = factor(rlgatnd_rev, levels = rev(0:7)))) +
  geom_bar(stat = "identity", position = "stack") +
  coord_flip() +
  facet_wrap(~year) +
  scale_fill_manual(
    values = c(
      "0" = "#00002E", "1" = "#002060", "2" = "#003090", "3" = "#4A90E2",
      "4" = "#8AB4F8", "5" = "#DDDDDD",
      "6" = "#FFE399", "7" = "#FFBE5C"
    ),
    name = "Frequency"
  ) +
  labs(
    title = "Attendance of Religious Services apart from Special Occasions",
    x = NULL,
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

-   Much higher share of Polish respondents attends religious services
    more often
-   Slovenia is somewhat evenly distributed
-   Germany has the highest shares of respondnets that never attend
    religious services throughout all the years

### by Country over Time on Average

```{r}
dt_filtered %>%
  filter(!is.na(rlgatnd_rev), !is.na(year)) %>%
  group_by(cntry, year) %>%
  summarise(mean_rlgatnd_rev = mean(rlgatnd_rev, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_rlgatnd_rev, color = cntry, group = cntry)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Attendance of Religious Services apart from Special Occasions",
    x = "Year",
    y = "Average Frequency (0–7 scale)",
    color = "Country"
  ) +
  theme_minimal()
```

-   Germany on average the country with the least attendance, followed
    by Slovenia and then Poland
-   Poland is by far the country with the highest attendance, even
    though there is an upward trend that peaked in 2020


# Models

## Correlation Matrix

```{r}
# Select relevant variables
dt_cor <- dt_filtered %>% select(trstep, ppltrst, euftf, stfedu, stfgov, stfhlth, trstlgl, imwbcnt, rlgatnd_rev, gndr, age_group, year) 

# Calculate correlation matrix (pairwise complete cases)
cor(dt_cor, use = "pairwise.complete.obs")
```

## Baseline Linear Model with all Variables as Predictors

```{r}
options(scipen = 999)

model1 <- lm(trstep ~ ppltrst + euftf + stfedu + stfgov + stfhlth + trstlgl + imwbcnt + rlgatnd_rev + gndr + year + age_group + cntry,
             data = dt_filtered)
summary(model1)

```

### Interpretation - Ab hier weiter in Interpretationen

**Model Overview**

-   **Dependent variable**: `trstep` – trust in the EU (0–10 scale)
-   **N** = 37,297 (after removing 17,244 rows with missing values)
-   **R² = 0.364** – The model explains \~36% of the variation in EU
    trust, which is quite good for social science research
-   **Residual SE = 1.98** – The typical prediction error is about 2
    points on a 0–10 scale
-   **Country controls (`cntry`)** are included, with Germany as the
    reference group
-   All predictors are statistically significant at p \< 0.001 except
    **`rlgatnd`**, which is **not significant**.

*Institutional and Political Attitudes*

-   **`trstlgl` (+0.29)**: Trust in the legal system is the **strongest
    predictor** of EU trust. A one-unit increase is associated with a
    **0.29-point increase** in EU trust.
-   **`euftf` (+0.21)**: Stronger support for European unification
    correlates with significantly higher trust in the EU.
-   **`stfgov` (+0.15)**: Satisfaction with national government
    substantially boosts trust in the EU.
-   **`stfedu` (+0.06)** and **`stfhlth` (+0.04)**: Satisfaction with
    education and healthcare modestly increase EU trust.
-   **`ppltrst` (+0.05)**: Greater interpersonal trust is positively
    associated with EU trust.
-   **`freehms` (−0.10)**: Belief that homosexuals are not free in
    society slightly **reduces** EU trust (higher scores = more
    disagreement with the statement that homosexuals are free).
-   **`imwbcnt` (+0.07)**: Positive attitudes toward immigrants are
    associated with increased EU trust.

*Demographics*

-   **`gndr` (+0.29)**: Being female (if coded as 1) is linked to
    **higher** trust in the EU.
-   **`age_group` (−0.13)**: Older age groups tend to trust the EU
    **less**.
-   **`rlgatnd` (ns)**: Religious attendance is **not significantly**
    related to EU trust.

*Time Trend*

-   **`year` (−0.019)**: Trust in the EU has slightly **declined** over
    time — about **0.019 points per year**.

*Country Differences* (relative to Germany)

-   **`cntryPL` (+0.72)**: Respondents in **Poland** report
    **significantly higher** trust in the EU compared to Germans.
-   **`cntrySI` (+0.77)**: Respondents in **Slovenia** also have
    **higher** trust in the EU than Germans.

*People who trust others, trust their legal system, support further EU
unification, and are satisfied with their national institutions
(especially government) tend to trust the EU more. Trust is higher in
Poland and Slovenia than Germany, and has declined slightly over time.*

##Separate Models per Country for Comparison This avoids running
regressions with interactions, that give long outputs that are hard to
interpret.


### Baseline Linear Model for Poland

```{r}
model_PL <- lm(trstep, ppltrst, euftf, stfedu, stfgov, stfhlth, trstlgl, imwbcnt, rlgatnd_rev, gndr, agea,
               data = filter(dt_filtered, cntry == "PL"))

summary(model_PL)
```

-   stfedu are not statistically significant
-   trstlgl, euftf, clsprty and gender are strongest predictors

### Baseline Linear Model for Slovenia

```{r}
model_SI <- lm(trstep ~ ppltrst + euftf + freehms + stfedu + stfgov + stfhlth + trstlgl + imwbcnt + rlgatnd + gndr + year + clsprty + age_group, 
               data = filter(dt_filtered, cntry == "SI"))

summary(model_SI)
```

-imwbcnt not statistically significant -trstlgl, gndr, clsprty stongest
predictors

### Baseline Linear Model for Germany

```{r}
model_DE <- lm(trstep, ppltrst, euftf, stfedu, stfgov, stfhlth, trstlgl, imwbcnt, rlgatnd_rev, gndr, agea, 
               data = filter(dt_filtered, cntry == "DE"))

summary(model_DE)
```

-year not statistically significant -stfgov, trstlgl, gndr strongest
predictors

## Interaction with Country for Comparative Analysis - Hier weiter im Code, der geht bei mir nicht

-This tests whether the effect of a variable differs by country. -The
model shows if any predictor significantly interacts with country. -This
means the model allows each predictor to have different effects in
Poland and Slovenia, compared to the reference country Germany
(baseline).

```{r}
model2 <- lm(trstep ~ (ppltrst, euftf, stfedu, stfgov, stfhlth, trstlgl, imwbcnt, rlgatnd_rev, gndr, agea, cntry) * cntry,
             data = dt_filtered)

summary(model2)
```

*Summary of Main Points* Key Takeaways - Legal system trust, EU
optimism, and satisfaction with national institutions drive EU trust in
all three countries, but the strength of these predictors varies.

-   Voting behavior, religiosity, and gender show important
    country-specific dynamics:

-   More religious people trust the EU more in Poland, but less in
    Germany and Slovenia.

-   Gender effect reverses: women trust more in Germany, less in
    Slovenia.

-   Age trends reverse: older Germans trust less, older Poles trust
    more.

-   Time trend: slight decline in EU trust over time in Poland and
    Slovenia, but no significant change in Germany.
    
## Single Interaction Models
Since we do not expect the effect of all predictors to be the same for all groups, we will test for different interactions.

### Political Orientation x Party Identification 
Do Party affilitation and EU-Skepticism together have an efffect?

```{r}
model3 <- lm(trstep ~ clsprty * euftf,
             data = dt_filtered)

summary(model3)
```

### Gender or Age Group x Institutional Trust
What role does the Gender or Age Group play together with institutional trust?

```{r}
model4 <- lm(trstep ~ age_group * stfhlth,
             data = dt_filtered)

summary(model4)
```

```{r}
model5 <- lm(trstep ~ gndr * trstlgl,
             data = dt_filtered)

summary(model5)
```


## Country-Specific Group Models 

- for comparison, because regressions with interactions give long outputs are hard to interpret

### Baseline Linear Model for Poland

```{r}

model_PL <- lm(trstep, ppltrst, euftf, stfedu, stfgov, stfhlth, trstlgl, imwbcnt, rlgatnd_rev, gndr, agea,
               data = filter(dt_filtered, cntry == "PL"))

summary(model_PL)
```

-   stfedu are not statistically significant
-   trstlgl, euftf, clsprty and gender are strongest predictors

### Baseline Linear Model for Slovenia

```{r}

model_SI <- lm(trstep ~ ppltrst + euftf + freehms + stfedu + stfgov + stfhlth + trstlgl + imwbcnt + rlgatnd + gndr + year + clsprty + age_group, 
               data = filter(dt_filtered, cntry == "SI"))

summary(model_SI)
```

-imwbcnt not statistically significant -trstlgl, gndr, clsprty stongest
predictors

### Baseline Linear Model for Germany

```{r}

model_DE <- lm(trstep, ppltrst, euftf, stfedu, stfgov, stfhlth, trstlgl, imwbcnt, rlgatnd_rev, gndr, agea, 
               data = filter(dt_filtered, cntry == "DE"))

summary(model_DE)
```

-year not statistically significant -stfgov, trstlgl, gndr strongest
predictors


## Diagnostics & Model Assumptions

*Multivariate Normality*
- analysis assumes that the residuals (differences between predicted and observed values) are normally distributed
- assumption can be assessed by examining histograms or Q-Q plots of the residuals, or through statistical tests such as Kolmogorov-Smirnov test

*No multicollienarity*
- correlation matrices, coefficients ideally below 0.8
- Variance Inflation Factor (VIF): values above 10 indicating problematic multicollinearity

*Homoscedasticity*
- the variance of error terms (residuals) should be consistent across all levels of the independent variables
- a scatterplot of residuals versus predicted values should not display and discernible pattern, such as a cone-shaped distribution 
- adressing it might involve data transformation or adding a quadratic term to the model
