---
title: "Project Study"
author: "Helena Zappe"
date: "2025-05-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Download Packages

```{r install-packages, eval=FALSE, include=FALSE}
install.packages(c(
  "tidyverse",  # Datenaufbereitung & Visualisierung
  "skimr",      # Übersichtliche Daten-Zusammenfassung
  "psych",      # Deskriptive Statistik
  "rmarkdown",  # Für R Markdown selbst
  "knitr",      # Code-Ausgabe in R Markdown
  "tinytex",    # Nur für PDF-Export nötig
  "labelled"
  ))
tinytex::install_tinytex()  # Nur beim ersten Mal notwendig
```

# Load Packages

```{r load-packages, include=FALSE}
library(tidyverse)   # Bündelt ggplot2, dplyr, readr, etc. – für Datenimport, -aufbereitung, Visualisierung
library(skimr)       # Für übersichtliche Zusammenfassungen von Datensätzen
library(psych)       # Für deskriptive Statistiken (z.B. Mittelwert, SD, Median)
library(rmarkdown)   # Für das Erstellen von HTML/PDF-Dokumenten
library(knitr)       # Macht Code-Ausgabe in R Markdown hübsch und steuerbar
library(tinytex)     # Wird nur gebraucht, wenn du PDF-Dateien erstellst
library(labelled)
library(dplyr)
```

# Loading data

```{r}

setwd("/Users/test/Documents/Studium/Semester 6/Project Study")

dt_raw <- haven::read_sav(file.choose())

```

# Variables

## Research Question: "What shapes trust in the European Union and United Nations in Poland, Germany, and Slovenia? A comparative analysis of political attitudes, ideology, and socio-demographics."

```{r}

polintr        # Political interest
imwbcnt        # Immigration attitude (higher = better view)
lrscale        # Ideology
stfdem         # Satisfaction with democracy
stfgov         # Satisfaction with government
eduyrs         # Education level
trstplt        # Trust in politicians
age_group      # Moderator
cntry          # Countries   
nwspol         # News about politics

trstun         # Trust in UN
trstep         # Trust in EU
```

# Data Preparation

```{r}
#filtering to our countries and recoding age group and reversing the scale for polintr (higher number = more interested)
dt_new <- dt_raw %>%
  filter(cntry %in% c("DE", "PL", "SI")) %>%
  mutate(
    age_group = case_when(
      agea < 18 ~ "<18",
      agea >=18 & agea <= 24 ~ "18-24",
      agea > 24 & agea <= 34 ~ "25-34",
      agea > 34 & agea <= 44 ~ "35-44",
      agea > 44 & agea <= 54 ~ "45–54",
      agea > 54 & agea <= 64 ~ "55-64",
      agea > 64 ~ "65+"
    ))  %>%
  mutate(
    age_group = factor(age_group, levels = c("<18", "18-24", "25-34", "35-44", "45–54", "55-64", "65+")),
    polintr_rev = 5 - polintr  # higher = more interested
    )

table(dt_new$age_group, useNA = "always")
table(dt_new$cntry, useNA = "always")

```

# Descriptive Statistics

```{r}

# trust in EU and UN over time for all three countries (flow chart)
# predictors for each country
# age groups per country (young or old country)
# satisfaction with government and democracy for each country

#Summary statistics of variables by age group
dt_new %>%
  group_by(age_group) %>%
  summarise(across(
    c(polintr_rev, eduyrs, lrscale, stfdem, stfgov, imwbcnt, polintr, trstplt, trstun, trstep),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  ))

#declining trust with age

dt_new %>%
  group_by(cntry) %>%
  summarise(across(
    c(trstplt, trstun, trstep),
    list(mean = ~mean(.x, na.rm = TRUE),
         median = ~median(.x, na.rm = TRUE),
         Q5 = ~quantile(.x, 0.05, na.rm = TRUE),
         Q95 = ~quantile(.x, 0.95, na.rm = TRUE))
  ))

#big differences across countries
```

# Visualization

```{r}

#Plotting average political interest vs. age group
dt_new %>%
  group_by(age_group) %>%
  summarise(mean_interest = mean(polintr_rev, na.rm = TRUE)) %>%
  ggplot(aes(x = age_group, y = mean_interest, fill = age_group)) +
  geom_col() +
  labs(title = "Average Political Interest by Age Group", y = "Political Interest")

#Plotting average trust in EU vs. age group
dt_new %>%
  group_by(age_group) %>%
  summarise(mean_trust = mean(trstep, na.rm = TRUE)) %>%
  ggplot(aes(x = age_group, y = mean_trust, fill = age_group)) +
  geom_col() +
  labs(title = "Average Trust in EU by Age Group", y = "Trust in EU")

#Plotting average trust in UN vs. age group
dt_new %>%
  group_by(age_group) %>%
  summarise(mean_trust = mean(trstun, na.rm = TRUE)) %>%
  ggplot(aes(x = age_group, y = mean_trust, fill = age_group)) +
  geom_col() +
  labs(title = "Average Trust in UN by Age Group", y = "Trust in UN")


  ## Compute mean trust in UN per country
  country_means <- dt_new %>%
    group_by(cntry) %>%
    summarise(mean_trstun = mean(trstun, na.rm = TRUE))
  
  # Plot
  ggplot(country_means, aes(x = cntry, y = mean_trstun)) +
    geom_point(size = 3, color = "steelblue") +
    labs(title = "Mean Trust in UN by Country", y = "Mean Trust in UN", x = "Country") +
    theme_minimal()
  

## Compute means per country and age group
group_means <- dt_new %>%
  group_by(cntry, age_group) %>%
  summarise(mean_trstun = mean(trstun, na.rm = TRUE), .groups = "drop")

# Plot
ggplot(group_means, aes(x = cntry, y = mean_trstun, color = age_group)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  labs(title = "Mean Trust in UN by Country and Age Group",
       y = "Mean Trust in UN", x = "Country") +
  theme_minimal()

#big differences among countries and age groups within the countries

#Correlation matrix
# Select relevant variables
dt_subset <- dt_new %>% select(trstun, trstep, polintr, imwbcnt, lrscale, stfdem, stfgov, eduyrs, trstplt, agea) #I use agea because age_group is a fator variable which I can not use for correlation matrix

# Calculate correlation matrix (pairwise complete cases)
cor(dt_subset, use = "pairwise.complete.obs")


```

# Models
## Trust in UN

```{r}

options(scipen = 999) #to get whole numbers in the output

#1. Age group only
mod1 <- lm(trstep ~ age_group, data = dt_new)
summary(mod1)

#2. Adding trstplt as control variable
mod2 <- lm(trstep ~ age_group + trstplt, data = dt_new)
summary(mod2)

#3. Interaction with trust to see if trust matters more/less by age
mod3 <- lm(trstep ~ age_group + trstplt + age_group * trstplt , data = dt_new)
summary(mod3)

#4. adding predictors
mod4 <- lm(trstep ~ age_group + trstplt + polintr + imwbcnt + lrscale + stfdem + stfgov + eduyrs, data = dt_new)
summary(mod4)

#5. Full interaction model: adding age*media interaction to see if the effect of news consumption differs by age group
mod5 <- lm(trstep ~ age_group * (trstplt + polintr + imwbcnt + lrscale + stfdem + stfgov + eduyrs), data = dt_new)
summary(mod5)

#7. interaction between media*trust to explore whether the impact of trust on political interest depends on media usage patterns.
mod6 <- lm(polintr_rev ~ age_group + trstplt * (nwspol + netustm + netusoft), data = dt_new)
summary(mod6)

#8. Fully Saturated Model (all Interactions) but only if we have enough data and want to explore complex interactions
mod7 <- lm(polintr_rev ~ age_group * (nwspol + netustm + netusoft) +
                            trstplt * (nwspol + netustm + netusoft), data = dt_raw)
summary(mod7)



```

## Trust in EU

# Implications / Predictions

