---
title: "Shiny Workshop"
author: "Marei Göbelbecker"
date: "2025-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Shiny Workshop

Look at the ones from last year, he mentioned:
https://sebastian-weibels.shinyapps.io/Populism-Elections/
https://sebastian-weibels.shinyapps.io/Sustainability-Elections/

-	Menu on top
-	Plots with reactive programming dependent on the filters applied
-	Hover for information in the plots

```{r, eval=FALSE, include=FAlSE}
install.packages("shiny")
library(shiny)
# library(tidyverse)
```

# Creating the App

```{r}
# Creating the App
ui <- fluidPage(
  "Hello world!"
)
  # here as fluid Page (scrollable, use this for smartphone use also, grid of 12)
  # otherwise also dashboard possible (rather wide format)

# Defining the function
server <- function(input, output, session) {
}

##  Opening the App Page
shinyApp(ui, server)
```

# Creating the user Interface

```{r}
ui <- fluidPage(
  titlePanel("Title Panel"),
  
  sidebarLayout(
    sidebarPanel("Title: SIdebar Panel"), 
  
  mainPanel("Title: Main Panel",
            fluidRow(
              column(width = 4, ...)
            ))
  
  )
)

# Defining the function
server <- function(input, output) {
}

##  Opening the App Page
shinyApp(ui = ui, server = server)
```

# Adding UI controls

- selecting data set as input for output 
- different inputs go into the ui selectInput 
- goes into server funtion 
- tableOutput then gives the output on the server funtion

**see markdown for special commands on inputs widgets**

### Exercises on inout functions

```{r}
# install.packages("shinyWidgets")
# library(shinyWidgets)

?textInput #use this for information on the functions

ui <- fluidPage(
  # 1: writing text within 
  textInput("Your Name", "", 
            placeholder = "What is your name?"),

  # 2: slider input 0-100, interval 5 with animation
  sliderInput("number", "Select a number", 
              min=0, max=100, 
              value=0, 
              step=5, 
              animate=TRUE),
  
  # 3: subheadlines in a list
  selectInput("breed", "Your favorite animal breed:", 
              choices = 
                list(dogs= list("Bulldog", "Labrador"),
                     cats= list("Persian cat", "Senegal cat"))
             ),
)

# remember the commas between the different Inputs!!!!

server <- function(input, output) {
}


shinyApp(ui, server)

```

# Output functions

create a placeholder functions which are then given out by the server function

### Text Output
```{r}
ui <- fluidPage(
  textOutput("text"),
  verbatimTextOutput("code")
)

server <- function(input, output, session) {
  output$text <- renderText({"Hello Fliend!"})
}

shinyApp(ui, server)
```

**look at his markdown for Table & Plot Output etc.**

# Exercise on Output
```{r}
?textOutput

# 1: Answer b

```

# Regression
R markdown for code

```{r}
install.packages("stargazer")
library(stargazer)
```

# Responsive Coding

When typing in an input, it is shown in the output. 
R markdown for Code

Render functions are by default reactive ->
in his code the difference is not in the output but in the function; useful if the engn string is used for more than one output, so we can use the same input on more than one part

the order of function definition & function calling doesn't matter

observer is not necessary but observer makes it cleaner: seperate into reactive and observer to make the dashboard and prozessorleistung easier :)

## building an app around the plot
4 = width (3*4=12 grid)
Input1=size
Input2=mean
Input3=sd

! need to use reactive ! normal function would not work

# Notes:
- think of a structure of the dashboard first
- coding top-down
- think about what is shown and which outputs should be interactive
- in which ones does it make sense to show it interactive? which filters are useful, which are not?

- only use the cleaned and refined dataset into the server function (one server for the full ui)
- possible structure: First by page, then  by reactive & observer, make comments!

