---
title: 'Data Science Week 1: Intro to R and R Markdown'
author: "Tom Zimmermann"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(repos = list(CRAN="http://cran.rstudio.com/"))
```


In this section, we want to take a first look at the Vienna hotels dataset introduced in the lecture. 

```{r}
# Housekeeping
#install.packages("tidyverse")
#install.packages("rmarkdown")
#install.packages("knitr")
#install.packages("data.table")

# Set working directory (in case it's needed)
# setwd("")
```

# Loading data

## Reading a csv file

`csv` files are a simple way to import data into different software. Read the `W1_HotelsVienna.csv` file using three different methods: Use the standard `read.csv` function, the `read_csv` function from the `readr` package and the `fread` function from the `data.table` package.

```{r}
# Insert code here

library(readr)
library(data.table)

dt1 <- read.csv(file = "W1_HotelsVienna.csv")
dt2 <- read_csv(file = "W1_HotelsVienna.csv")
dt3 <- fread(file = "W1_HotelsVienna.csv")

# End
```


## Reading a foreign file

Sometimes files are only available in a software-specific format. Read the `W1_HotelsVienna.dta` file using the `read_dta` function from the `haven` package.

```{r}
# Insert code here

library(haven)

dt4 <- read_dta(file = "W1_HotelsVienna.dta")

# End
```


## Reading a file directly from the internet

Instead of downloading and saving a file locally first, you can also read files from the internet directly. For example, the URL for the `W1_HotelsVienna.csv` dataset is https://osf.io/y6jvb/download . Import the file directly using `read_csv`.


```{r}
# Insert code here

dt5 <- read_csv("https://osf.io/y6jvb/download")

# End
```


# Data structures

Load the Vienna hotels data with any method you like into an object called `dt`.

```{r}
# 
dt <- read_csv("https://osf.io/y6jvb/download")

# End
```


0. Does the table contain cross-sectional or time-series data?

Cross-sectional data: many different hotels (units) observed at the same time (weekday Nov 2017)

1. What is the data type of the `neighbourhood` column in the data?

```{r}
# Insert code here

class(dt$neighbourhood)

# End
```

2. The `str()` function is helpful to get a quick overview of the different variable values and variable types. How many variables of type `character` are in the data?

```{r}
# Insert code here

str(dt)
table(sapply(dt, class))

# End
```

3. Add a variable `Number total ratings` as the sum of the `rating_count` and `ratingta_count` variables.

```{r}
# Insert code here

dt$number_total_ratings <- (dt$rating_count + dt$ratingta_count)

# End
```

4. Create a vector named `offers` with the first 10 elements of the `offer_cat` variable and another vector named `city` with the first element of the `city_actual` variable.

```{r}
# Insert code here

offers <- dt$offer_cat[1:10]
city   <- dt$city_actual[1]

# End
```

5. Combine your vectors `offers` and `city` into a list.

```{r}
# Insert code here

my_list <- list(offers = offers, city = city)
my_list

# End
```





