---
title: 'Data Science Week 2: Data Wrangling with Tidyverse'
author: "Tom Zimmermann"
output:
  html_document:
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```


In this section, we want to practice data wrangling with `tidyverse` functions. We use the World Bank immunization dataset discussed in the lecture.
You can download the data from the course website or from the following url: https://osf.io/gk5cn/download


1. Read the data with `read_csv`.

```{r}
# Insert code here

worldBank <- read_csv(url("https://osf.io/gk5cn/download"))

# End
```



2. Create a new table with only the columns `year`, `countryname`, `imm` and `pop`.


```{r}
# Insert code here

worldBank2 <- worldBank %>% 
  select(year, countryname, imm, pop)

# End
```

3. Population is measured in millions. Add a new variable to `worldBank2` that measures population in 1000s.

```{r}
# Insert code here

worldBank2 <- worldBank2 %>% 
  mutate(pop_1000 = pop*1000)

# End
```


4. Calculate the mean and median immunization rates over all countries from your table `worldBank2`.

```{r}
# Insert code here

worldBank2 %>% 
  summarise(mean_imm   = mean(imm), 
            median_imm = median(imm))

# End
```


5. Calculate the mean and median immunization rates over all countries *for each year separately*.

```{r}
# Insert code here

worldBank2 %>% 
  group_by(year) %>% 
  summarise(across(imm, 
                   list(mean   = mean, 
                        median = median)))

# End
```


6. Create two tables that both include `year` and `countryname`. One table (call it `df_pop`) includes population as only additional variable. One table (`df_imm`) includes the measle immunization rate as only additional variable.

```{r}
# Insert code here

df_pop <- worldBank %>% 
  select(year, countryname, pop)
df_imm <- worldBank %>% 
  select(year, countryname, imm)

# End
```


7. Change `df_imm` such that it does only include observations from the year 2017.

```{r}
# Insert code here

df_imm <- df_imm %>% 
  filter(year == 2017)

# End
```

8. Join `df_imm` onto `df_pop` using a `left_join`. Try again and use an `inner_join`. Note how the results differ.

```{r}
# Insert code here

df_joined_1 <- left_join(df_pop, df_imm)
df_joined_2 <- inner_join(df_pop, df_imm)

# With left_join all rows in x (so df_pop) are included. With inner_join, only rows that show up in both x and y (so in df_pop and df_imm) are included.

# End
```
