---
title: 'Data Science Week 4: Exploratory Data Analysis and Data Visualization'
author: "Tom Zimmermann"
output:
  html_document:
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```


In this section, we look into exploratory data analysis and data visualization using the World Bank immunization dataset discussed in previous lectures.

You can download the data from the course website or from the following url: https://osf.io/gk5cn/download



0. Import the dataset into R

```{r}
# Insert code here

dt = read_csv("https://osf.io/gk5cn/download")

# End
```


1. Calculate the mean, median, the 5th and 95th quantile of the immunization rate variable (`imm` in the dataset) over non-missing values.

```{r}
# Insert code here

# Here are a couple of built-in functions that let you compute these quantities
mean(dt$imm, na.rm = TRUE)
median(dt$imm, na.rm = TRUE)
quantile(dt$imm, .05, na.rm = TRUE)
quantile(dt$imm, .95, na.rm = TRUE)

# Can you use these functions within dplyr? This would make life easier for some of the next exercises.
dt %>% 
  summarise(meanIMM   = mean(imm, na.rm = TRUE),
            medianIMM = median(imm, na.rm = TRUE),
            Q5IMM     = quantile(imm, .05, na.rm = TRUE),
            Q95IMM    = quantile(imm, .95, na.rm = TRUE))

# End
```



3. Calculate different measures of spread of `imm`: Calculate the range, the interquartile range and the standard deviation of `imm` *in 2017*.

```{r}
# Insert code here
dt %>% 
  filter(year == 2017) %>% 
  summarise(range = max(imm, na.rm = TRUE) -  min(imm, na.rm = TRUE),
            IQR   = quantile(imm, .75, na.rm = TRUE) - quantile(imm, .25, na.rm = TRUE),
            SD    = sd(imm, na.rm = TRUE))

# End
```

3. Calculate mean-median skewness of `imm` for each year in the data separately using the formula in the slides.

```{r}
# Insert code here
dt %>% 
  group_by(year) %>% 
  summarise(Skew = (mean(imm, na.rm = TRUE) - median(imm, na.rm = TRUE))/sd(imm, na.rm = TRUE))

# End
```


4. Repeat question 1 but compute summary statistics for each year separately. Has the mean immunization rate increased or decreased over time?

```{r}
# Insert code here

dt %>% 
  group_by(year) %>% 
  summarise(meanIMM   = mean(imm, na.rm = TRUE),
            medianIMM = median(imm, na.rm = TRUE),
            Q5IMM     = quantile(imm, .05, na.rm = TRUE),
            Q95IMM    = quantile(imm, .95, na.rm = TRUE))

# End
```

5. Plot histograms of `gdppc` and `imm` in 2017. Do any observations look strange to you? Also explore the effect of the number of bins.

```{r}
# Insert code here

#gddpc
dt %>% filter(year == 2017) %>% 
  ggplot(aes(x = gdppc)) +
  geom_histogram(binwidth = 2000)

dt %>% filter(year == 2017) %>% 
  ggplot(aes(x = gdppc)) +
  geom_histogram(binwidth = 100)

#imm
dt %>% 
  filter(year == 2017) %>% 
  ggplot(aes(imm)) +
  geom_histogram(binwidth = 1)

dt %>% 
  filter(year == 2017) %>% 
  ggplot(aes(imm)) +
  geom_histogram(binwidth = 10)

# End
```


6. Use your results from question 4 to plot mean immunization rates over time.

```{r}
# Insert code here

dt %>% 
  group_by(year) %>% 
  summarise(meanIMM = mean(imm, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = meanIMM)) +
  geom_point() +
  geom_line()


# Alternatively
dt %>% 
  ggplot(aes(x = year, y = imm)) +
  stat_summary_bin()  # Calculates mean and se for each year by default

# End
```

6. Create a scatter plot with `gdppc` on the x-axis and `imm` on the y-axis to explore the relationship between these two variables in 2017. (It might be useful to plot `gdppc` on a log scale)


```{r}
# Insert code here
dt %>% 
  filter(year == 2017) %>% 
  ggplot(aes(x = gdppc, y = imm)) +
  geom_point() +
  scale_x_log10()

# End
```

7. Extend your plot from the previous part to include data from the years 2000 and 2017. Use different colors for each year.

```{r}
# Insert code here

dt %>% 
  filter(year == 2000 | year == 2017) %>% 
  ggplot(aes(x = gdppc, y = imm, color = as.factor(year))) +
  geom_point() +
  scale_x_log10()

# End
```

8. Prettify the plot in part 7 by e.g. adding proper axis labels, title, caption or legend.

```{r}
# Insert code here

dt %>% 
  filter(year == 2000 | year == 2017) %>% 
  ggplot(aes(x = gdppc, y = imm, color = as.factor(year))) +
  geom_point() +
  scale_x_log10(labels = scales::dollar) +
  theme_minimal(base_size = 18) +
  labs(x = "GDP Per Capita",
       y = "Immunization rate",
       title = "GDP and Immunization Rates",
       subtitle = "Data points are country-years",
       caption = "Source: World Bank.",
       color = '')

# End
```

