---
title: 'Data Science Week 5: Simple linear regression'
author: "Tom Zimmermann"
output:
  html_document:
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
includeSolutions = FALSE
library(tidyverse)
```



In this section, we look at simple linear regression using the World Bank life expectancy dataset from DABEP.
You can download the data from the course website or from the following url: https://osf.io/sh9mu/download


1. Load the data and keep only observations from the year 2017.

```{r}
# Insert code here

df_worldbank <- read_csv("https://osf.io/sh9mu/download") %>% 
  filter(year == 2017)

# End
```


2. Make a scatter plot of `gdppc` (x-axis) against `lifeexp` (y-axis). Add a linear regression line. Do you think this is a good fit?


```{r}
# Insert code here
df_worldbank %>% 
  ggplot(aes(x = gdppc, y = lifeexp)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, formula = y~x)



# End
```

3. Redo the plot with regression line but now use the natural log of `gdppc` as x variable. What do you think about the fit now?

```{r}
# Insert code here
df_worldbank %>% 
  ggplot(aes(x = log(gdppc), y = lifeexp)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, formula = y~x)


# End
```


4. Estimate the linear regression $lifeexp^E = \alpha + \beta ln(gdppc)$ and provide results. How do you interpret $\hat{\beta}$?

```{r}
# Insert code here

fitLM1 <- lm(formula = lifeexp ~ log(gdppc), 
             data = df_worldbank)

summary(fitLM1)

# A 1% increase in GDP p.c. is associated with an increase in life expectancy of 0.053336 years on average. 

# End
```

4. Estimate the linear regression $lifeexp^E = \alpha + \beta gdppc$ (no log). How do you interpret $\hat{\beta}$? Does this model provide a better or worse fit than the model in part 3 based on comparing their R$^2$s?

```{r}
# Insert code here

fitLM2 <- lm(formula = lifeexp ~ gdppc, 
             data = df_worldbank)

summary(fitLM2)

# End
```

5. Expand the regression in part 4 by adding a quadratic term in `gdppc`, that is, estimate the model $lifeexp^E = \alpha + \beta_1 gdppc + \beta_2 gdppc^2$. Do the estimates suggest an inverted U-shaped relationship? Does this model provide a better or worse fit than the model in part 4 based on comparing their R$^2$s?

```{r}
# Insert code here

fitLM3 <- lm(formula = lifeexp ~ gdppc + I(gdppc^2), 
             data = df_worldbank)

summary(fitLM3)

# End
```

6. Plot the quadratic function estimated in part 5 together with the scatter and the linear model from part 2.


```{r}
# Insert code here
df_worldbank %>% 
  ggplot(aes(x = gdppc, y = lifeexp)) +
  geom_point() + 
  geom_smooth(method = lm, formula = y~x, se = FALSE, colour = 'blue') +
  geom_smooth(method = lm, formula = y ~ x + I(x^2), se = FALSE, color = 'orange') +
  theme_bw()


# End
```
